---
title: "Nonlinear models with natural spline"
author: "Yanelli Nunez"
date: "12/15/2021"
output:
 html_document:
    toc: TRUE
    toc_float: TRUE
    code_folding: hide
editor_options: 
  chunk_output_type: console
---

Note: Some of the code chunks take a while to run

*Objective:*  

To run the models for the main analysis. 

I save the gam model output and relationship curve in a dataset so that I can use it later to make figures etc. So here I am running the models, plotting the models, and saving the results. Here, I already know which relationships are linear and which are not. 

#Economic models:
Include one economic variable per model

#Hispanic models:
Include % Hispanic as the exposure

#Asian and American Native:
One model per race variable adjusted for % White using s(perc_white)

#Black and White:
These models include Black and White variables in the model. If nonlinear, the variable was included with a natural spline, with 4 degrees of freedom and reference to the mean


Notes: 
-Relationships that deviate from linearity are included with a natural spline and 4 degrees of freedom with reference to the mean

-Four variables were included in all models to adjust for potential confounding:
1) Urbanicity - as a categorical variable
2) Year - as a categorical variable
3) EPA region - as categorical  
4) Population density - continuous with a penalized spline


```{r include=FALSE}

rm(list=ls())

# 1a Declare root directory
project.folder <- paste0(print(here::here()),'/')

# add folder locations
source(paste0(project.folder,'0_00_create_folder_structure.R'))

# add file locations
source(paste0(file.locations.folder,'external_file_locations.R'))

# load packages
source(paste0(packages.folder,'packages_to_load.R'))

# load functions
source(paste0(code.folder, 'models/functions.R')) ## All functions I used in this .rmd are located in this script

# load data
source(paste0(code.folder, 'data_loading.R'))

```

## i. View Data Summaries 
```{r}
summary(data_master)
summary(so2_energy)
```


## 1. Prep function arguments

I created a dataset with the info for the model functions' arguments (see functions.R script), which I then split into the following subdatasets:

Nonlinear_models: 
1) economic variables --> all economic variables have a non-linear relationship with the outcome

2) racial variables --> most racial variables are non-linear
    2a. Asian, Hispanic, and American Native models are all adjusted for s(perc_white)
    2b. Black and White are model together by adjusting the model with ns(perc_white, df = 4)

Linear_models:

1) racial variables --> only racial variables have linear associations
    1a. Asian, Hispanic, and American Native models are all adjusted for s(perc_white)
    1b. Black model adjusted for ns(perc_white, df = 4)
    1c. Single Hispanic model

With these datasets, I can then just loop the model functions I made across the dataset rows to run all of the models    
    
Notes: 
- I had to split the nonlinear models into economic and racial because the racial models are adjusted for percent white but the economic models are not, so I couldn't run them in the same batch with the workflow i have.
- I also run the models for percent black and percent white separately because in this case we include both variables with a natural spline (for the other racial models we adjust using s(perc_white))

```{r}

### dataset containing the arguments for all models we need to run

models_to_make = c("Ns(perc_unemp, df = 4, ref = 5.4)",  # we use the mean as the reference for the nonlinear curves 
             "Ns(perc_pvert, df = 4, ref = 16.45)",
             "Ns(fam_incom_adjs, df = 4, ref = 55047)",
             "Ns(propty_val_adjs, df = 4, ref = 93712)",
             "Ns(perc_white, df = 4, ref = 87.58)",
             "Ns(perc_black, df = 4, ref = 8.81)",
             "Ns(perc_asian, df = 4, ref = 1.08)",
             "Ns(perc_nativ, df = 4, ref = 1.29)",
             "Ns(perc_hisp_org, df = 4, ref = 4.42)") %>%
  as.data.frame() %>%
   rename(., ns_arg = .) %>%
  mutate(exposure =  c("perc_unemp", 
                  "perc_pvert",
                  "fam_incom_adjs",
                  "propty_val_adjs",
                  "perc_white",
                  "perc_black",
                  "perc_asian",
                  "perc_nativ",
                  "perc_hisp_org")) %>%
  as.data.frame() %>%
  mutate(x_label = c("% Unemployment", "% Poverty", "Median Family Income ($1000)", "Median Propty Value ($1000)", "% White", "% Black", "% Asian", "% American Native", "% Hispanic")) %>% # x label for ggplot
  mutate(color = dplyr::if_else(exposure %in% c("perc_unemp", "perc_pvert", "fam_incom_adjs", "propty_val_adjs"), "#69b3a2", "#386cb0")) %>% # specify ggplot color based on exposure variable
 mutate(freq=7) %>%
  map_df(., rep, .$freq) %>%
mutate(id = rep(1:7, length.out = n())) %>%
mutate(outcome = case_when(id == 1 ~ 'so2_energy_relat_chg',  ## add outcome column
                           id == 2 ~ 'nox_energy_relat_chg', 
                           id == 3 ~ 'so2_indus_relat_chg', 
                           id == 4 ~ 'nox_trans_relat_chg', 
                           id == 5 ~ 'nh3_agric_relat_chg', 
                           id == 6 ~ 'oc_resid_relat_chg', 
                           id == 7 ~ 'nox_commer_relat_chg')) %>%
  dplyr::select(-freq, -id) %>%
   mutate(data_spf = if_else(outcome == "so2_energy_relat_chg", "so2_energy",
                      if_else(outcome == "so2_indus_relat_chg", "so2_inds",
                        if_else(outcome == "nox_energy_relat_chg", "nox_energy", "data_master")))) %>%
  arrange(outcome)
  

### dataset with the arguments for the NON-LINEAR MODELS  
  models_to_make_nonlinear <- models_to_make %>%
  dplyr::filter(!(outcome == "so2_energy_relat_chg" & exposure %in% c("perc_asian", "perc_hisp_org"))) %>%
  dplyr::filter(!(outcome == "nox_energy_relat_chg" & exposure %in% c("perc_white", "perc_black", "perc_asian", "perc_nativ"))) %>%
  dplyr::filter(!(outcome == "so2_indus_relat_chg" & exposure == "perc_black")) %>%
  dplyr::filter(!(outcome == "nox_trans_relat_chg" & exposure == "perc_nativ")) %>%
  dplyr::filter(!(outcome == "nh3_agric_relat_chg" & exposure == "perc_nativ")) %>%
  dplyr::filter(!(outcome == "oc_resid_relat_chg" & exposure == "perc_black")) 
  
  
### dataset with arguments for the NON-LINEAR models: economic variables & hispanic
  models_to_make_nonlinear_eco <- models_to_make_nonlinear %>%
    filter(exposure %in% c("perc_unemp", "perc_pvert", "fam_incom_adjs", "propty_val_adjs", "perc_hisp_org")) 
    

### dataset with arguments for the NON-LINEAR models: asian and american native (adjusted for perc_white)
### 1) includes asian, native, and hispanic 
  models_to_make_nonlinear_race <- models_to_make_nonlinear %>%
    filter(exposure %in% c("perc_asian", "perc_nativ"))
  
## 2) includes Black 
  models_to_make_nonlinear_blk_wht <- models_to_make_nonlinear %>%
    filter(exposure == "perc_black")
  
  
### dataset with the arguments for the LINEAR MODELS 
  models_to_make_linear <- models_to_make %>%
  dplyr::filter(outcome == "so2_energy_relat_chg" & exposure %in% c("perc_asian", "perc_hisp_org") |
                outcome == "nox_energy_relat_chg" & exposure %in% c("perc_white", "perc_black", "perc_asian", "perc_nativ") |
                outcome == "so2_indus_relat_chg" & exposure == "perc_black" |
                outcome == "nox_trans_relat_chg" & exposure == "perc_nativ" |
                outcome == "nh3_agric_relat_chg" & exposure == "perc_nativ" |
                outcome == "oc_resid_relat_chg" & exposure == "perc_black") %>%
    mutate(basis_name = case_when(exposure == "perc_white" & data_spf == "nox_energy" ~ 'cb_nox_energy_white',
                             exposure == "perc_black" & data_spf == "nox_energy" ~ 'cb_nox_energy_black',
                             exposure == "perc_black" & data_spf == "so2_inds" ~ 'cb_so2_inds_black',
                             exposure == "perc_black" & data_spf == "data_master" ~ 'cb_data_master_black',
                             exposure == "perc_asian" & data_spf == "so2_energy" ~ 'cb_so2_energy_asian',
                             exposure == "perc_asian" & data_spf == "nox_energy" ~ 'cb_nox_energy_asian',
                             exposure == "perc_nativ" & data_spf == "nox_energy" ~ 'cb_nox_energy_nativ',
                             exposure == "perc_nativ" & data_spf == "data_master" ~ 'cb_data_master_nativ',
                             exposure == "perc_hisp_org" & data_spf == "so2_energy" ~ 'cb_so2_energy_hisp')) %>%
    mutate(pred_name = gsub("cb", "pred", basis_name)) %>%
    mutate(pred_name = if_else(outcome == "nox_trans_relat_chg", "pred_nox_trans_nativ", pred_name))
  
  
  ### dataset with arguments for the LINEAR racial models
### 1) includes asian, native, and hispanic 
  models_to_make_linear_race <- models_to_make_linear %>%
    filter(exposure %in% c("perc_asian", "perc_nativ"))
  
## 2) includes Black 
  models_to_make_linear_blk_wht <- models_to_make_linear %>%
    filter(exposure == "perc_black",
           outcome != "nox_energy_relat_chg") # because in this case white is also linear 

```


## 2. Progress bar 

Parallelize: runs the iterations in parallel across multiple cores to increase speed
You computer has multiple processing cores but it only uses one to run your R code. Parallelizing the processing means that you are telling your computer to run different portions of your code simultaneously in multiple cores. This results in overall faster processing

Code for Progress bar and paralleling of point 3a.

```{r}

#### set up a progress bar
pb <- txtProgressBar(min = 0, max = nrow(models_to_make_nonlinear_eco), width = 25, style = 3)
progress <- function(p) setTxtProgressBar(pb, p)
opts <- list(progress=progress)

### code to parallelize the processing among multiple cores
cores <- ceiling(parallel::detectCores() / 2) # use half of the available cores 
cluster <- snow::makeCluster(cores) 
doSNOW::registerDoSNOW(cluster)

```

## 3. Autoprocessing of NON-LINEAR models

- We use the foreach() function to loop our function over each row of the dataset 'models_to_make_nonlinear_eco'.

### 3a. Nonlinear economic and hispanic models

These models are NOT adjusted for % white
The code chunk takes some time to run

```{r}

### run the foreach() loop (I am parallelizing this process to make it faster)
         
models_nonlin_eco <- foreach(i = iterators::icount(nrow(models_to_make_nonlinear_eco)), # iterate across rows
                             .options.snow=opts, # the progress bar being sent to each core
                             .packages=c("Epi", "magrittr", "dplyr", "janitor", "ggplot2"), # the packages being sent to each core
                             .export = ls(globalenv()),
                             .inorder = FALSE, # speeds up parallel programming
                             .combine = rbind) %dopar% { # if you doing parallel programming use %dopar% otherwise use %do%
  
  # make model i
  m <- run_model(data   = models_to_make_nonlinear_eco$data_spf[[i]],
                outcome = models_to_make_nonlinear_eco$outcome[[i]], 
               exposure = models_to_make_nonlinear_eco$ns_arg[[i]]) # the natural spline terms
  
  df <- models_to_make_nonlinear_eco[i,] %>% mutate(models = list(m))
  
  # make plot for model i (don't need to index by i since that was done in df above^)
 plot <- plot_nonlin(model = df$models[[1]], 
                     data = df$data_spf[[1]], 
                 exposure = df$exposure[[1]],
                   x_label = df$x_label[[1]],
                     color = df$color[[1]])
 
 df <- df %>% mutate(plots = list(plot)) %>%
  dplyr::select(-ns_arg, -color, -x_label) %>%
  arrange(outcome)
 
  
  # save our work
  df
  
}

# stop the parellel processing
snow::stopCluster(cluster)
close(pb)


# save list dataset 
saveRDS(models_nonlin_eco, '/Users/jennishearston/Desktop/YN/results_updated/models_nonlin_eco_hisp.rds')
#saveRDS(models_nonlin_eco, paste0(yanelli_external, "results_updated/models_nonlin_eco_hisp.rds"))

```


### 3a. Nonlinear racial models

#### Asian & American Native 

These models are adjusted for s(perc_white)
The code chunk takes some time to run

```{r}

#### set up a progress bar
pb <- txtProgressBar(min = 0, max = nrow(models_to_make_nonlinear_race), width = 25, style = 3)
progress <- function(p) setTxtProgressBar(pb, p)
opts <- list(progress=progress)

### code to paralize the processing among multiple cores
cores <- ceiling(parallel::detectCores() / 2) # use half of the available cores 
cluster <- snow::makeCluster(cores) 
doSNOW::registerDoSNOW(cluster)

### run models
         
models_nonlin_race <- foreach(i = iterators::icount(nrow(models_to_make_nonlinear_race)), # iterate across rows
                             .options.snow=opts, # the progress bar being sent to each core
                             .packages=c("Epi", "magrittr", "dplyr", "janitor", "ggplot2"), # the packages being sent to each core
                             .export = ls(globalenv()),
                             .inorder = FALSE, # speeds up parallel programming
                             .combine = rbind) %dopar% { # if you doing parallel programming use %dopar% otherwise use %do%
  
  # make model i
  m <- run_model_race(data   = models_to_make_nonlinear_race$data_spf[[i]],
                     outcome = models_to_make_nonlinear_race$outcome[[i]], 
                    exposure = models_to_make_nonlinear_race$ns_arg[[i]]) # the natural spline terms
  
  df <- models_to_make_nonlinear_race[i,] %>% mutate(models = list(m))
  
  # make plot for model i (don't need to index by i since that was done in df above^)
 plot <- plot_nonlin(model = df$models[[1]], 
                     data = df$data_spf[[1]], 
                 exposure = df$exposure[[1]],
                   x_label = df$x_label[[1]],
                     color = df$color[[1]])
 
 df <- df %>% mutate(plots = list(plot)) %>%
  dplyr::select(-ns_arg, -color, -x_label) %>%
   arrange(outcome)
  
  # save our work
  df
  
}


# stop the parellel processing
snow::stopCluster(cluster)
close(pb)


# save list dataset 
saveRDS(models_nonlin_race, '/Users/jennishearston/Desktop/YN/results_updated/models_nonlin_asian_an.rds')
#saveRDS(models_nonlin_race, paste0(yanelli_external, "results_updated/models_nonlin_asian_an.rds"))

```


#### Black & White 

These models include black and white race percentage as nonlinear terms (natural splines)
The code chunk takes some time to run

```{r}

#### set up a progress bar
pb <- txtProgressBar(min = 0, max = nrow(models_to_make_nonlinear_blk_wht), width = 25, style = 3)
progress <- function(p) setTxtProgressBar(pb, p)
opts <- list(progress=progress)

### code to paralized the processing among multiple cores
cores <- ceiling(parallel::detectCores() / 2) # use half of the available cores 
cluster <- snow::makeCluster(cores) 
doSNOW::registerDoSNOW(cluster)

### run the models using a foreach() loop
         
models_nonlin_blk_wht <- foreach(i = iterators::icount(nrow(models_to_make_nonlinear_blk_wht)), # iterate across rows
                             .options.snow=opts, # the progress bar being sent to each core
                             .packages=c("Epi", "magrittr", "dplyr", "janitor", "ggplot2"), # the packages being sent to each core
                             .export = ls(globalenv()),
                             .inorder = FALSE, # speeds up parallel programming
                             .combine = rbind) %dopar% { # if you doing parallel programming use %dopar% otherwise use %do%
  
  # make model i
  m <- run_model_blk(data   = models_to_make_nonlinear_blk_wht$data_spf[[i]],
                     outcome = models_to_make_nonlinear_blk_wht$outcome[[i]], 
                    exposure = models_to_make_nonlinear_blk_wht$ns_arg[[i]]) # the natural spline terms
  
  df <- models_to_make_nonlinear_blk_wht[i,] %>% mutate(models = list(m))
  
  # make plot for model i (don't need to index by i since that was done in df above^)
 plot <- plot_nonlin(model = df$models[[1]], 
                     data = df$data_spf[[1]], 
                 exposure = df$exposure[[1]],
                   x_label = df$x_label[[1]],
                     color = df$color[[1]])
 
 df <- df %>% mutate(plots = list(plot)) 
 
  # make plot for White for model i (don't need to index by i since that was done in df above^)
 plot_white <- plot_nonlin_white(model = df$models[[1]], 
                     data = df$data_spf[[1]], 
                     color = df$color[[1]])
 
 df <- df %>% mutate(plots_white = list(plot_white)) %>%
  dplyr::select(-ns_arg, -color, -x_label)
  
  # save our work
  df
  
}

# stop the parellel processing
snow::stopCluster(cluster)
close(pb)


# save list dataset 
saveRDS(models_nonlin_blk_wht, '/Users/jennishearston/Desktop/YN/results_updated/models_nonlin_blk_wht.rds')
# saveRDS(models_nonlin_blk_wht, paste0(yanelli_external, "results_updated/models_nonlin_blk_wht.rds"))

```

## 4. Autoprocessing of LINEAR models

 The associations below are linear:

1) SO2 from energy: 
-Asian
-Hispanic

2) NOx from energy:
-White
-Black
-Asian
-A. Native

3) SO2 from industry:
-Black

4) NOx from transport 
-American native 

5) NH3 from agriculture 
-American native

6) OC from residential
-Black

7) NOx from commercial
-all nonlinear terms

### 4a. Onebasis

To plot the linear models in the same style as the nonlinear ones (using a reference to the mean), I need to use a onebasis for the independent variable. Here is the code to create the one basis for all linear models. 

```{r}
### run the foreach() loop to generate the onebasis 
         
foreach(i = iterators::icount(nrow(models_to_make_linear)), # iterate across rows
                             .combine = rbind) %do% { 
  
  # make basis
  b <- one_basis(data = models_to_make_linear$data_spf[[i]],
               exposure = models_to_make_linear$exposure[[i]])
  
   assign(models_to_make_linear$basis_name[[i]], b) 
   
  
}

remove(b) ## this is basically the last i which is also saved with a specific name in the environment


```


### Models

#### Asian, American Native

These models are adjusted for s(perc_white)

```{r}

#### set up a progress bar
pb <- txtProgressBar(min = 0, max = nrow(models_to_make_nonlinear_blk_wht), width = 25, style = 3)
progress <- function(p) setTxtProgressBar(pb, p)
opts <- list(progress=progress)

## Run models using a foreach() loop

linear_models_race <- foreach(i = iterators::icount(nrow(models_to_make_nonlinear_blk_wht)), # iterate across rows
                             .options.snow=opts, # the progress bar 
                             .inorder = FALSE, 
                             .combine = rbind) %do% { 
  
  # make models
      m <- run_model_race(data = models_to_make_linear_race$data_spf[[i]],
                       outcome = models_to_make_linear_race$outcome[[i]],
                      exposure = models_to_make_linear_race$basis_name[[i]]) ## in this case use the onebasis as the exposure 
  
   df <- models_to_make_linear_race[i,] %>% mutate(models = list(m))
   
df

}

remove(m)


```

#### Hispanic 

Only the model for SO2 from Energy is linear for hispanics. This model is NOT adjusted for % White because % White includes hispanic and nonhispanic white

```{r}

hisp_linear_model <- run_model(data = models_to_make_linear$data_spf[[9]],
                            outcome = models_to_make_linear$outcome[[9]], 
                           exposure = models_to_make_linear$basis_name[[9]])

summary(hisp_linear_model)

hisp_linear_model_df <- models_to_make_linear[9,] %>% mutate(models = list(hisp_linear_model))

linear_models_race <- rbind(hisp_linear_model_df, linear_models_race) %>%
  arrange(outcome)


```


#### Black & White 

These two models include both black and white. Black is included it as linear term and White included as nonlinear term with a natural spline with 4df 

```{r}

#### set up a progress bar
pb <- txtProgressBar(min = 0, max = nrow(models_to_make_nonlinear_blk_wht), width = 25, style = 3)
progress <- function(p) setTxtProgressBar(pb, p)
opts <- list(progress=progress)

### Run models

linear_models_blk_wht <- foreach(i = iterators::icount(nrow(models_to_make_linear_blk_wht)), # iterate across rows
                             .options.snow=opts, # the progress bar being sent to each core
                             .inorder = FALSE, # speeds up parallel programming
                             .combine = rbind) %do% { 
  
  # make models
      m <- run_model_blk(data = models_to_make_linear_blk_wht$data_spf[[i]],
                      outcome = models_to_make_linear_blk_wht$outcome[[i]],
                     exposure = models_to_make_linear_blk_wht$basis_name[[i]]) ## in this case use the onebasis as the exposure
  
   df <- models_to_make_linear_blk_wht[i,] %>% mutate(models = list(m))
   
   
df

}

```

#### Black and white linear

For NOx from the Energy sector, both Black and White are linear so I'm including the terms here linearly 

```{r}

nox_energy_lin_race = gamm4::gamm4(nox_energy_relat_chg ~ cb_nox_energy_black +
                                            cb_nox_energy_white +
                                            epa_reg +
                                            s(pop_density) +
                                            urbanicity +
                                            year, 
                         random = ~(1|state/gisjoin), data = nox_energy) 

# Bind all linear models


black_nox <- models_to_make_linear[3,] %>% mutate(models = list(nox_energy_lin_race[["gam"]]))
white_nox <- models_to_make_linear[2,] %>% mutate(models = list(nox_energy_lin_race[["gam"]]))

nox_blk_white <- rbind(black_nox, white_nox) %>%
  arrange(outcome)

linear_models <- rbind(linear_models_blk_wht, linear_models_race) %>%
  rbind(., nox_blk_white) %>%
  arrange(outcome) 


```

### 4b. Crosspred

I couldn't make crosspred work by referencing the basis so I couldn't run the crosspred portion inside foreach(). Thus, I have to do it here individually for each model :( (total of 10)

```{r}

pred_nox_energy_white <- crosspred(
  #basis = get(models_to_make_linear$basis_name[1]), #### the basis argument isn't working using indexing :(
  basis = cb_nox_energy_white, #### the basis argument isn't working using indexing :(
  model = linear_models$models[[5]], 
  at = get(linear_models$data_spf[[5]])[, linear_models$exposure[[5]]], 
  cen = (mean(get(linear_models$data_spf[[5]])[, linear_models$exposure[[5]]])))

pred_nox_energy_black <- crosspred(
  basis = cb_nox_energy_black, #### the basis argument isn't working using indexing :(
  model = linear_models$models[[4]], 
  at = get(linear_models$data_spf[[4]])[, linear_models$exposure[[4]]], 
  cen = (mean(get(linear_models$data_spf[[4]])[, linear_models$exposure[[4]]])))

pred_so2_inds_black <- crosspred(
  basis = cb_so2_inds_black, #### the basis argument isn't working using indexing :(
  model = linear_models$models[[10]], 
  at = get(linear_models$data_spf[[10]])[, linear_models$exposure[[10]]], 
  cen = (mean(get(linear_models$data_spf[[10]])[, linear_models$exposure[[10]]])))

pred_data_master_black <- crosspred(
  basis = cb_data_master_black, #### the basis argument isn't working using indexing :(
  model = linear_models$models[[7]], 
  at = get(linear_models$data_spf[[7]])[, linear_models$exposure[[7]]], 
  cen = (mean(get(linear_models$data_spf[[7]])[, linear_models$exposure[[7]]])))

pred_so2_energy_asian <- crosspred(
  basis = cb_so2_energy_asian, #### the basis argument isn't working using indexing :(
  model = linear_models$models[[9]], 
  at = get(linear_models$data_spf[[9]])[, linear_models$exposure[[9]]], 
  cen = (mean(get(linear_models$data_spf[[9]])[, linear_models$exposure[[9]]])))

pred_nox_energy_asian <- crosspred(
  basis = cb_nox_energy_asian, #### the basis argument isn't working using indexing :(
  model = linear_models$models[[2]], 
  at = get(linear_models$data_spf[[2]])[, linear_models$exposure[[2]]], 
  cen = (mean(get(linear_models$data_spf[[2]])[, linear_models$exposure[[2]]])))

pred_nox_energy_nativ <- crosspred(
  basis = cb_nox_energy_nativ, #### the basis argument isn't working using indexing :(
  model = linear_models$models[[3]], 
  at = get(linear_models$data_spf[[3]])[, linear_models$exposure[[3]]], 
  cen = (mean(get(linear_models$data_spf[[3]])[, linear_models$exposure[[3]]])))

pred_nox_trans_nativ <- crosspred(
  basis = cb_data_master_nativ, #### the basis argument isn't working using indexing :(
  model = linear_models$models[[6]], 
  at = get(linear_models$data_spf[[6]])[, linear_models$exposure[[6]]], 
  cen = (mean(get(linear_models$data_spf[[6]])[, linear_models$exposure[[6]]])))

pred_data_master_nativ <- crosspred(
  basis = cb_data_master_nativ, #### the basis argument isn't working using indexing :(
  model = linear_models$models[[1]], 
  at = get(linear_models$data_spf[[1]])[, linear_models$exposure[[1]]], 
  cen = (mean(get(linear_models$data_spf[[1]])[, linear_models$exposure[[1]]])))

pred_so2_energy_hisp <- crosspred(
  basis = cb_so2_energy_hisp, #### the basis argument isn't working using indexing :(
  model = linear_models$models[[8]], 
  at = get(linear_models$data_spf[[8]])[, linear_models$exposure[[8]]], 
  cen = (mean(get(linear_models$data_spf[[8]])[, linear_models$exposure[[8]]])))

```

### 4c. Plot linear models 

Here I am making the plots for the linear models
This code chunk takes a bit to run

```{r}

#### set up a progress bar
pb <- txtProgressBar(min = 0, max = nrow(linear_models), width = 25, style = 3)
progress <- function(p) setTxtProgressBar(pb, p)
opts <- list(progress=progress)

### code to paralized the processing among multiple cores
cores <- ceiling(parallel::detectCores() / 2) # use half of the available cores 
cluster <- snow::makeCluster(cores) 
doSNOW::registerDoSNOW(cluster)


## Run models

linear_plots <- foreach(i = iterators::icount(nrow(linear_models)), # iterate across rows
                             .options.snow=opts, # the progress bar being sent to each core
                             .packages=c("Epi", "magrittr", "dplyr", "janitor", "ggplot2"), # packages being sent to each core
                             .export = ls(globalenv()),
                             .inorder = FALSE, # speeds up parallel programming
                             .combine = rbind) %dopar% { 
  
  # crosspred 
      ##pred <- crosspred(
      #              basis = linear_models$basis_name[[i]], #### the basis argument isn't working using indexing :(
      #              model = linear_models$models[[i]], 
      #                 at = get(linear_models$data_spf[[i]])[, linear_models$exposure[[i]]], 
      #                cen = (mean(get(linear_models$data_spf[[i]])[, linear_models$exposure[[i]]])))                         
     
      #df <- linear_models[i,] %>% mutate(predic_est = list(pred))
      
  # plot models                            
      p <- plot_linear(pred_est = linear_models$pred_name[[i]],
                        x_label = linear_models$x_label[[i]],
                          color = linear_models$color[[i]]) ## in this case use the onebasis as the exposure
  
  
  df <- linear_models[i,] %>% mutate(plots = list(p)) %>%
        dplyr::select(-ns_arg, -color, -x_label, -basis_name, -pred_name)
   
   df
}


# stop the parellel processing
snow::stopCluster(cluster)
close(pb)

```


## 5. Save file
No need to run this code
```{r}

# saveRDS(linear_plots, paste0(yanelli_external, "results_updated/models_linear.rds"))

#linear_plots <- readRDS(paste0(yanelli_external, "results_updated/models_linear.rds"))

```


## 6. Plot nonlinear white

In the case of organic carbon from residential and so2 from industry, the association with Black is linear and with White nonlinear. In the code chunck above, I plot the linear association for Black and in this code chunk, using the same models, I plot the nonlinear associations for White

```{r}

nonl_white <- linear_plots %>%
  filter(outcome %in% c("oc_resid_relat_chg", "so2_indus_relat_chg")) %>%
  dplyr::select(-plots) %>%
  mutate(exposure = "perc_white") %>%
  left_join(., models_to_make_nonlinear)
  

# plot models for white

nonl_white_plots <- foreach(i = iterators::icount(nrow(nonl_white)), # iterate across rows
                             .inorder = FALSE, # speeds up parallel programming
                             .combine = rbind) %do% { # if you doing parallel programing use %dopar% otherwise use %do%

# plot_white
                               
   plot_white <- plot_nonlin_white(model = nonl_white$models[[i]], 
                                    data = nonl_white$data_spf[[i]], 
                                   color = nonl_white$color[[i]])
 
# add plot to data
 
    df  <- nonl_white[i,] %>% mutate(plots_white = list(plot_white)) %>%
  dplyr::select(-ns_arg, -color, -x_label)
  
  # save our work
  df
  
 
}

# save plots

##saveRDS(nonl_white_plots, paste0(yanelli_external, "results_updated/models_nonlin_wht.rds"))

```

