---
title: "Regional Models"
author: "Yanelli Nunez"
date: "1/16/2022"
output:
 html_document:
    toc: TRUE
    toc_float: TRUE
    code_folding: hide
editor_options: 
  chunk_output_type: console
---

Objective:  

Run regional models.

We have a total of:

-8 EPA regions





The exposures I run are: 


```{r include=FALSE}

rm(list=ls())

# 1a Declare root directory
project.folder <- paste0(print(here::here()),'/')

# add folder locations
source(paste0(project.folder,'0_00_create_folder_structure.R'))

# add file locations
source(paste0(file.locations.folder,'external_file_locations.R'))

# load packages
source(paste0(packages.folder,'packages_to_load.R'))

# load functions
source(paste0(code.folder, 'models/functions_yanelli.R'))

# load data
source(paste0(code.folder, 'data_loading.R'))

```

## 1. Prep function arguments

This dataset includes information to run all models with nonlinear relationships

```{r}

# dataset containing the arguments for all models we need to run

models_to_make <-  c("perc_unemp", 
                  "perc_pvert",
                  "fam_incom_adjs",
                  "propty_val_adjs",
                  "perc_white",
                  "perc_black",
                  "perc_asian",
                  "perc_nativ",
                  "perc_hisp_org") %>%
  as.data.frame() %>%
  rename(., exposure = .) %>%
  mutate(x_label = c("% Unemployment", "% Poverty", "Median Family Income ($1000)", "Median Propty Value ($1000)", "% White", "% Black", "% Asian", "% American Native", "% Hispanic")) %>% # x label for ggplot
  mutate(color = dplyr::if_else(exposure %in% c("perc_unempl", "perc_pvert", "fam_incom_adjs", "propty_val_adjs"), "#69b3a2", "#386cb0")) %>% # specify ggplot color based on exposure variable
 mutate(freq=7) %>%
  map_df(., rep, .$freq) %>%
mutate(id = rep(1:7, length.out = n())) %>%
mutate(outcome = case_when(id == 1 ~ 'so2_energy_relat_chg',  ## add outcome column
                           id == 2 ~ 'nox_energy_relat_chg', 
                           id == 3 ~ 'so2_indus_relat_chg', 
                           id == 4 ~ 'nox_trans_relat_chg', 
                           id == 5 ~ 'nh3_agric_relat_chg', 
                           id == 6 ~ 'oc_resid_relat_chg', 
                           id == 7 ~ 'nox_commer_relat_chg')) %>%
  dplyr::select(-freq, -id) %>%
   mutate(data_spf = if_else(outcome == "so2_energy_relat_chg", "so2_energy",
                      if_else(outcome == "so2_indus_relat_chg", "so2_inds",
                        if_else(outcome == "nox_energy_relat_chg", "nox_energy", "data_master")))) %>%    mutate(freq = 8) %>%
  map_df(., rep, .$freq) %>%
mutate(id = rep(1:8, length.out = n())) %>%
mutate(epa_reg = case_when(id == 1 ~ 'region_1_3',  ## add outcome column
                           id == 2 ~ 'region_4', 
                           id == 3 ~ 'region_5', 
                           id == 4 ~ 'region_6', 
                           id == 5 ~ 'region_7', 
                           id == 6 ~ 'region_8', 
                           id == 7 ~ 'region_9',
                           id == 8 ~ 'region_10')) %>%
   dplyr::select(-freq, -id) %>%
  filter(exposure %in% c("fam_incom_adjs", "perc_black", "perc_white"),
         outcome %in% c("nox_energy_relat_chg", "so2_indus_relat_chg", "so2_energy_relat_chg")) ## I am  running  only  a couple of the regionals
  
```

## 2. Progress bar 

Parallelize: runs the iterations in parallel across multiple cores to increase speed
You computer has multiple processing cores but it only uses one to ran your R code. Paralizing the procesing means that you are telling your computer to run different portions of your code simultaneously in multiple cores. This results in an overall faster processing

```{r}

#### set up a progress bar
pb <- txtProgressBar(min = 0, max = nrow(models_to_make), width = 25, style = 3)
progress <- function(p) setTxtProgressBar(pb, p)
opts <- list(progress=progress)

### code to paralized the processing among multiple cores
cores <- ceiling(parallel::detectCores() / 2) # use half of the available cores 
cluster <- snow::makeCluster(cores) 
doSNOW::registerDoSNOW(cluster)

```


## 3. autoprocessing of REGIONAL models

```{r}

### run the foreach() loop ( I am parallelizing this process to make it faster)
         
models_regional <- foreach(i = iterators::icount(nrow(models_to_make)), # iterate across rows
                             .options.snow=opts, # the progress bar being sent to each core
                             .packages=c("Epi", "magrittr", "dplyr", "janitor", "ggplot2"), # the packages being sent to each core
                             .export = ls(globalenv()),
                             .inorder = FALSE, # speeds up parallel programming
                             .combine = rbind) %dopar% { # if you doing parallel programing use %dopar% otherwise use %do%
# make model i

 m <- reg_model_func(data = models_to_make$data_spf[[i]],
                     epa_region = models_to_make$epa_reg[[i]],
                        outcome = models_to_make$outcome[[i]], 
                       exposure = models_to_make$exposure[[i]])

# add models to df
df <- models_to_make[i,] %>% mutate(models = list(m))

# make plots  
 p <- plot_nonlin_regional(model = df$models[[1]],
            data = df$data_spf[[1]],
        exposure = df$exposure[[1]],
          region = df$epa_reg[[1]],
           color = df$color[[1]],
         x_label = df$x_label[[1]])
   
# add plots to df
df <- df %>% mutate(plots = list(p)) %>%
  dplyr::select(-color, -x_label) 
  
df

}

# stop the parellel processing
snow::stopCluster(cluster)
close(pb)
```

## 4. Arrange dataset

```{r}
models_regional <- models_regional %>%
  mutate(epa_reg = fct_relevel(epa_reg, "region_1_3", "region_4", "region_5", "region_6", "region_7", "region_8", "region_9", "region_10")) %>%
  arrange(outcome, exposure, epa_reg)
```

## 5. Explore results

### 5a. NOx from energy

```{r}

pdf(file = "energy_nox_reg.pdf", width = 11, height = 13)            
 grid.arrange(models_regional$plots[[1]],
              models_regional$plots[[2]],
              models_regional$plots[[3]],
              models_regional$plots[[4]],
              models_regional$plots[[5]],
              models_regional$plots[[6]],
              models_regional$plots[[7]],
              models_regional$plots[[8]],
              
              # Perc  Black
              models_regional$plots[[9]],
              models_regional$plots[[10]],
              models_regional$plots[[11]],
              models_regional$plots[[12]],
              models_regional$plots[[13]],
              models_regional$plots[[14]],
              models_regional$plots[[15]],
              models_regional$plots[[16]],
              
              # Perc White
              models_regional$plots[[17]],
              models_regional$plots[[18]],
              models_regional$plots[[19]],
              models_regional$plots[[20]],
              models_regional$plots[[21]],
              models_regional$plots[[22]],
              models_regional$plots[[23]],
              models_regional$plots[[24]],
             
             nrow = 6, ncol = 4,
             left = gridtext::richtext_grob("Energy: NO<sub>x</sub> (% chg)", rot=90,
             gp=gpar(fontsize=15,font=8)))
dev.off() 


```

### 5b. SO2 from energy

```{r}

pdf(file = "energy_so2_reg.pdf", width = 11, height = 13)            
 grid.arrange(models_regional$plots[[25]],
              models_regional$plots[[26]],
              models_regional$plots[[27]],
              models_regional$plots[[28]],
              models_regional$plots[[29]],
              models_regional$plots[[30]],
              models_regional$plots[[31]],
              models_regional$plots[[32]],
              
              # Perc  Black
              models_regional$plots[[33]],
              models_regional$plots[[34]],
              models_regional$plots[[35]],
              models_regional$plots[[36]],
              models_regional$plots[[37]],
              models_regional$plots[[38]],
              models_regional$plots[[39]],
              models_regional$plots[[40]],
              
              # Perc White
              models_regional$plots[[41]],
              models_regional$plots[[42]],
              models_regional$plots[[43]],
              models_regional$plots[[44]],
              models_regional$plots[[45]],
              models_regional$plots[[46]],
              models_regional$plots[[47]],
              models_regional$plots[[48]],
             
             nrow = 6, ncol = 4,
             left = gridtext::richtext_grob("Energy: SO<sub>2</sub> (% chg)", rot=90,
             gp=gpar(fontsize=15,font=8)))
dev.off() 
```

### 5b. SO2 from industry

```{r}

pdf(file = "indus_so2_reg.pdf", width = 11, height = 13)            
 grid.arrange(models_regional$plots[[49]],
              models_regional$plots[[50]],
              models_regional$plots[[51]],
              models_regional$plots[[52]],
              models_regional$plots[[53]],
              models_regional$plots[[54]],
              models_regional$plots[[55]],
              models_regional$plots[[56]],
              
              # Perc  Black
              models_regional$plots[[57]],
              models_regional$plots[[58]],
              models_regional$plots[[59]],
              models_regional$plots[[60]],
              models_regional$plots[[61]],
              models_regional$plots[[62]],
              models_regional$plots[[63]],
              models_regional$plots[[64]],
              
              # Perc White
              models_regional$plots[[65]],
              models_regional$plots[[66]],
              models_regional$plots[[67]],
              models_regional$plots[[68]],
              models_regional$plots[[69]],
              models_regional$plots[[70]],
              models_regional$plots[[71]],
              models_regional$plots[[72]],
             
             nrow = 6, ncol = 4,
             left = gridtext::richtext_grob("Industry: SO<sub>2</sub> (% chg)", rot=90,
             gp=gpar(fontsize=15,font=8)))
dev.off() 
```