---
title: "Regional Models"
author: "Yanelli Nunez"
date: "1/16/2022"
output:
 html_document:
    toc: TRUE
    toc_float: TRUE
    code_folding: hide
editor_options: 
  chunk_output_type: console
---

Objective:  

Run regional models for the following sectors:

Industry SO2
Transport NOx

We only evaluate 3 independent variables: 

Median family income
Black
White

We have a total of:

-8 EPA regions



```{r include=FALSE}

rm(list=ls())

# 1a Declare root directory
project.folder <- paste0(print(here::here()),'/')

# add folder locations
source(paste0(project.folder,'0_00_create_folder_structure.R'))

# add file locations
source(paste0(file.locations.folder,'external_file_locations.R'))

# load packages
source(paste0(packages.folder,'packages_to_load.R'))

# load functions
source(paste0(code.folder, 'models/functions.R'))

# load data
source(paste0(code.folder, 'data_loading.R'))

```

## 1. Prep function arguments

This dataset includes information to run all models with nonlinear relationships

```{r}

# dataset containing the arguments for all models we need to run

models_to_make <-  c("fam_incom_adjs",
                     "perc_black") %>%
  as.data.frame() %>%
  rename(., exposure = .) %>%
  mutate(x_label = c("Median Family Income ($1000)", "% Black")) %>% # x label for ggplot
  mutate(color = dplyr::if_else(exposure == "fam_incom_adjs", "#69b3a2", "#386cb0")) %>% # specify ggplot color based on exposure variable
 mutate(freq=7) %>%
  map_df(., rep, .$freq) %>%
mutate(id = rep(1:7, length.out = n())) %>%
mutate(outcome = case_when(id == 1 ~ 'so2_energy_relat_chg',  ## add outcome column
                           id == 2 ~ 'nox_energy_relat_chg', 
                           id == 3 ~ 'so2_indus_relat_chg', 
                           id == 4 ~ 'nox_trans_relat_chg', 
                           id == 5 ~ 'nh3_agric_relat_chg', 
                           id == 6 ~ 'oc_resid_relat_chg', 
                           id == 7 ~ 'nox_commer_relat_chg')) %>%
  dplyr::select(-freq, -id) %>%
   mutate(data_spf = if_else(outcome == "so2_energy_relat_chg", "so2_energy",
                      if_else(outcome == "so2_indus_relat_chg", "so2_inds",
                        if_else(outcome == "nox_energy_relat_chg", "nox_energy", "data_master")))) %>%    mutate(freq = 8) %>%
  map_df(., rep, .$freq) %>%
mutate(id = rep(1:8, length.out = n())) %>%
mutate(epa_reg = case_when(id == 1 ~ 'region_1_3',  ## add outcome column
                           id == 2 ~ 'region_4', 
                           id == 3 ~ 'region_5', 
                           id == 4 ~ 'region_6', 
                           id == 5 ~ 'region_7', 
                           id == 6 ~ 'region_8', 
                           id == 7 ~ 'region_9',
                           id == 8 ~ 'region_10')) %>%
   dplyr::select(-freq, -id) %>%
  mutate(epa_reg_name = case_when(epa_reg == 'region_1_3' ~ "Region 1-3",  ## add outcome column
                                  epa_reg == 'region_4' ~  "Region 4", 
                                  epa_reg == 'region_5' ~  "Region 5", 
                                  epa_reg == 'region_6' ~  "Region 6", 
                                  epa_reg == 'region_7' ~  "Region 7", 
                                  epa_reg == 'region_8' ~  "Region 8", 
                                  epa_reg == 'region_9' ~  "Region 9",
                                  epa_reg == 'region_10' ~ "Region 10")) %>%
  filter(outcome %in% c("so2_indus_relat_chg", "nox_trans_relat_chg")) ## I am  running  only  two sectors

# splitting the data into two sets, one for race and one for family income because the race model needs to include both Black and White together

models_to_make_race <- models_to_make %>%
  filter(exposure == "perc_black")

models_to_make_eco <- models_to_make %>%
  filter(exposure == "fam_incom_adjs")
```


## 2. Progress bar 

Parallelize: runs the iterations in parallel across multiple cores to increase speed
You computer has multiple processing cores but it only uses one to ran your R code. Paralizing the procesing means that you are telling your computer to run different portions of your code simultaneously in multiple cores. This results in an overall faster processing

```{r}

#### set up a progress bar
pb <- txtProgressBar(min = 0, max = nrow(models_to_make), width = 25, style = 3)
progress <- function(p) setTxtProgressBar(pb, p)
opts <- list(progress=progress)

### code to paralized the processing among multiple cores
cores <- ceiling(parallel::detectCores() / 2) # use half of the available cores 
cluster <- snow::makeCluster(cores) 
doSNOW::registerDoSNOW(cluster)

```


## 3. REGIONAL models median family income

```{r}

### run the foreach() loop ( I am paralleling this process to make it faster)
         
models_regional_eco <- foreach(i = iterators::icount(nrow(models_to_make_eco)), # iterate across rows
                             .options.snow=opts, # the progress bar being sent to each core
                             .packages=c("Epi", "magrittr", "dplyr", "janitor", "ggplot2"), # the packages being sent to each core
                             .export = ls(globalenv()),
                             .inorder = FALSE, # speeds up parallel programming
                             .combine = rbind) %dopar% { # if you doing parallel programming use %dopar% otherwise use %do%
# make model i

 m <- reg_model_func(data = models_to_make_eco$data_spf[[i]],
                     epa_region = models_to_make_eco$epa_reg[[i]],
                        outcome = models_to_make_eco$outcome[[i]], 
                       exposure = models_to_make_eco$exposure[[i]])

# add models to df
df <- models_to_make_eco[i,] %>% mutate(models = list(m))

# make plots  
 p <- plot_nonlin_regional(model = df$models[[1]],
            data = df$data_spf[[1]],
        exposure = df$exposure[[1]],
          region = df$epa_reg[[1]],
           color = df$color[[1]],
         x_label = df$x_label[[1]],
          title  = df$epa_reg_name[[1]])
   
# add plots to df
df <- df %>% mutate(plots = list(p)) %>%
  dplyr::select(-color, -x_label) 
  
df

}

# stop the parallel processing
snow::stopCluster(cluster)
close(pb)

### Arrange dataset

models_regional_eco <- models_regional_eco %>%
  mutate(epa_reg = fct_relevel(epa_reg, "region_1_3", "region_4", "region_5", "region_6", "region_7", "region_8", "region_9", "region_10")) %>%
  arrange(outcome, exposure, epa_reg)

### Save dataset
saveRDS(models_regional_eco, paste0(yanelli_external, "results_updated/sensit_analys_models_regional_eco.rds"))

```


## 4. REGIONAL models Black/White

```{r}

### run the foreach() loop ( I am paralleling this process to make it faster)
         
models_regional_race <- foreach(i = iterators::icount(nrow(models_to_make_race)), # iterate across rows
                             .options.snow=opts, # the progress bar being sent to each core
                             .packages=c("Epi", "magrittr", "dplyr", "janitor", "ggplot2"), # the packages being sent to each core
                             .export = ls(globalenv()),
                             .inorder = FALSE, # speeds up parallel programming
                             .combine = rbind) %dopar% { # if you doing parallel programming use %dopar% otherwise use %do%
# make model i

 m <- reg_model_func_race( data = models_to_make_race$data_spf[[i]],
                     epa_region = models_to_make_race$epa_reg[[i]],
                        outcome = models_to_make_race$outcome[[i]], 
                       exposure = models_to_make_race$exposure[[i]])

# add models to df
df <- models_to_make_race[i,] %>% mutate(models = list(m))

# make plots  
 p <- plot_nonlin_regional(model = df$models[[1]],
                            data = df$data_spf[[1]],
                        exposure = df$exposure[[1]],
                          region = df$epa_reg[[1]],
                           color = df$color[[1]],
                         x_label = df$x_label[[1]],
                          title  = df$epa_reg_name[[1]])
   
# add plots to df
df <- df %>% mutate(plots_black = list(p)) 
  
p_white <- plot_nonlin_regional_wt(model = df$models[[1]],
                                    data = df$data_spf[[1]],
                                  region = df$epa_reg[[1]],
                                   color = df$color[[1]],
                                  title  = df$epa_reg_name[[1]])

df <- df %>% mutate(plots_white = list(p_white)) %>%
  dplyr::select(-color, -x_label, -epa_reg_name) 

df

}

# stop the parallel processing
snow::stopCluster(cluster)
close(pb)

### Arrange dataset

models_regional_race <- models_regional_race %>%
  mutate(epa_reg = fct_relevel(epa_reg, "region_1_3", "region_4", "region_5", "region_6", "region_7", "region_8", "region_9", "region_10")) %>%
  arrange(outcome, exposure, epa_reg)

### Save dataset
saveRDS(models_regional_race, paste0(yanelli_external, "results_updated/sensit_analys_models_regional_race.rds"))

```



