---
title: "Correlation in exposures"
author: "Yanelli Nunez"
date: "7/10/2021"
output:
 html_document:
    toc: TRUE
    toc_float: TRUE
    code_folding: hide
editor_options: 
  chunk_output_type: console
---

Objective:  

Evaluate correlation between independent variables 

```{r include=FALSE}

rm(list=ls())

# 1a Declare root directory

project.folder <- paste0(print(here::here()),'/')

# add folder locations
source(paste0(project.folder,'0_00_create_folder_structure.R'))

# add file locations
source(paste0(file.locations.folder,'external_file_locations.R'))

# load packages
source(paste0(packages.folder,'packages_to_load.R'))

# load data
source(paste0(code.folder, 'data_loading.R'))

```


## 2. Scale data

```{r}

# scale
data_scaled <- data_master %>%
  mutate_at(c("perc_unemp", "perc_nohs", "perc_pvert", "perc_white", "perc_black", "perc_nativ", "perc_asian", "perc_hisp_org", "propty_val_adjs", "fam_incom_adjs", "pop_density"), ~(./sd(., na.rm = TRUE))) %>%
  arrange(year) %>%
  as.data.frame()

#class(data_scaled)

summary(data_scaled)
```

## 3. Viz

- Estimate correlations and plot 
- Percent Black is VERY correlated with percent White
- Percent Asian is positively correlated with percent Hispanic
- I double checked the data generation process and everything seems correct so the correlation results accurately represent our data 

```{r}

exp_var <- data_scaled %>%
  dplyr::select( pop_density, propty_val_adjs, fam_incom_adjs, perc_unemp, perc_pvert, perc_hisp_org, perc_nativ, perc_asian, perc_black, perc_white) %>%
  drop_na() 
  
corr_mat <- round(cor(exp_var, method = "pearson"), 2)
corr_mat

# corrplot::corrplot(corr_mat, method = "color") quick way of making a correlation plot 

corr_mat[upper.tri(corr_mat)] <- NA  # replace the upper righ corner with NA's

# melt correlation matrix
melted_cormat <- melt(corr_mat, na.rm = TRUE) %>%
  dplyr::rename(Correlation = value) %>%
  drop_na() 

#melted_cormat$X1 <- factor(melted_cormat$X1, levels = c("pop_density", "perc_white", "perc_black", "perc_asian", "perc_nativ", "perc_hisp_org", "perc_pvert", "perc_unemp", "fam_incom_adjs", "propty_val_adjs"))

#melted_cormat$X2 <- factor(melted_cormat$X2, levels = c("propty_val_adjs", "fam_incom_adjs", "perc_unemp", "perc_pvert", "perc_hisp_org", "perc_nativ", "perc_asian", "perc_white", "perc_black", "pop_density"))


melted_cormat$x_label <- factor(melted_cormat$X2, labels = c("Population~density", "Property~Value", "Family~Income", "'%'*Unemployment", "'%'*Poverty", "'%'*Hispanic", "'%'*American~Indian",  "'%'*Asian", "'%'*Black","'%'*White"))

melted_cormat %>%
    ggplot(aes(X1, X2, fill = Correlation)) +
geom_tile(color = "white") +
scale_fill_gradient2(low = "blue", high = "red", mid = "white",
                      midpoint = 0.5,  space = "Lab",
                      name = "Pearson\nCorrelation") +
  geom_text(aes(X1, X2, label = Correlation), color = "black", size = 6) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, vjust = 1, size = 18, hjust = 1),
    axis.text.y = element_text(vjust = 1, size = 18, hjust = 1),
  axis.title.x = element_blank(),
 axis.title.y = element_blank(),
 # panel.grid.major = element_blank(),
#  panel.border = element_blank(),
 # panel.background = element_blank(),
#  axis.ticks = element_blank(),
  legend.direction = "vertical",
  legend.text = element_text(size = 21),
  legend.title = element_text(size = 21)) +
# scale_y_discrete("X1", labels = c(expression(Pop~density), expression('%'*White), expression('%'*Black), expression('%'*Asian), expression('%'*American~Native), expression('%'*Hispanic), expression('%'*Poverty), expression('%'*Unemployment), expression(Family~Income), expression(Property~Value))) +
 scale_x_discrete("X2", labels = parse(text = levels(melted_cormat$x_label))) +
 scale_y_discrete("X1", labels = parse(text = levels(melted_cormat$x_label)))
```

## data viz
```{r}

 library(ggmosaic)

results <- read_xlsx(paste0(data.folder, "generated_data/summary_results.xlsx")) %>%
  gather(emissions, association, 3:9) %>%
  mutate(asso_num = case_when(association == "positive" ~ 1,
                              association == "negative" ~ -1,
                              association == "null" ~ 0))


results$ind_var <- factor(results$ind_var, levels = c("prop_val", "fam_income", "unempl", "poverty", "hispanic", "a_native", "asian", "black", "white"))
                          
results$label_x <- factor(results$ind_var, labels = c( "Property~Value", "Family~Income", "'%'*Unemployment", "'%'*Poverty", "'%'*Hispanic", "'%'*American~Native",  "'%'*Asian", "'%'*Black","'%'*White"))

 results %>%
  # filter(category == "race") %>%
            ggplot() +
 geom_col(aes(x = asso_num, y = ind_var, fill = association)) +
   facet_wrap(~emissions) +
   #facet_wrap(~exposure, ncol = 1, strip.position = "left") + 
   theme_minimal() +
   geom_vline(xintercept = 0) +
   theme(
    axis.text.x = element_text(angle = 45, vjust = 1, size = 12, hjust = 1))
                 
                #conds = product(rude_to_recline))) 

 results %>%
   ggplot() +
   geom_tile(aes(y = emissions, x = ind_var, fill = association)) +
   scale_fill_manual(values = c("#7fc97f", "#beaed4", "#fdc086"))

```

