---
title: "Dose response figures"
author: "Yanelli Nunez"
date: "1/28/2022"
output:
 html_document:
    toc: TRUE
    toc_float: TRUE
    code_folding: hide
editor_options: 
  chunk_output_type: console
---

Objective:  

Plot the dose-responses 


Note, you won't be able to run this code because the model outputs are save in my local drive (they are pretty big)

```{r include=FALSE}

rm(list=ls())

# 1a Declare root directory

project.folder <- paste0(print(here::here()),'/')

# add folder locations
source(paste0(project.folder,'0_00_create_folder_structure.R'))

# add file locations
source(paste0(file.locations.folder,'external_file_locations.R'))

# load packages
source(paste0(packages.folder,'packages_to_load.R'))

# load functions
 source(paste0(code.folder, 'models/functions.R'))

# load data
source(paste0(code.folder, 'data_loading.R'))

```


## 1. Load datasets 

These are the model result

```{r}

models_linear <- readRDS(paste0(yanelli_external, "results_updated/models_linear.rds")) %>%
  arrange(outcome)

models_nonlin_asian_an <- readRDS(paste0(yanelli_external, "results_updated/models_nonlin_asian_an.rds")) %>%
  arrange(outcome)

models_nonlin_blk_wht <- readRDS(paste0(yanelli_external, "results_updated/models_nonlin_blk_wht.rds")) %>%
  arrange(outcome)

models_nonlin_eco_hisp <- readRDS(paste0(yanelli_external, "results_updated/models_nonlin_eco_hisp.rds")) %>%
  arrange(outcome)

models_nonlin_wht <- readRDS(paste0(yanelli_external, "results_updated/models_nonlin_wht.rds")) %>%
  arrange(outcome)

```


### 1a. Combine datasets
Merging some of the datasets together to make the next steps easier
```{r}


models_nonlin_wht <- models_nonlin_wht %>%
  rename(plots = plots_white)

models_nonlinear <- rbind(models_nonlin_asian_an, models_nonlin_eco_hisp) %>%
  rbind(., models_nonlin_wht)

rm(models_nonlin_asian_an)
rm(models_nonlin_eco_hisp)
rm(models_nonlin_wht)
```


## 2. Extract plots into environment

```{r}

### nonlinear
nonlinear_rnm <- models_nonlinear %>%
  unite(name, outcome:exposure, sep = "_") %>%
  mutate(name = str_remove(name, "_relat_chg"))

# extract nonlinear using assign and foreach()
foreach(i = iterators::icount(nrow(nonlinear_rnm)), # iterate across rows
                             .combine = rbind) %do% { 
  
assign(nonlinear_rnm$name[[i]], nonlinear_rnm$plots[[i]])
                             }

### linear 
linear_rnm <- models_linear %>%
   unite(name, outcome:exposure, sep = "_") %>%
  mutate(name = str_remove(name, "_relat_chg"))


# extract nonlinear using assign and foreach()
foreach(i = iterators::icount(nrow(linear_rnm)), # iterate across rows
                             .combine = rbind) %do% { 
  
assign(linear_rnm$name[[i]], linear_rnm$plots[[i]])
                             }


### other white and black nonlinear models
nh3_agric_perc_white <- models_nonlin_blk_wht$plots_white[[1]]
nox_commer_perc_white <- models_nonlin_blk_wht$plots_white[[2]]
nox_trans_perc_white <- models_nonlin_blk_wht$plots_white[[3]]
so2_energy_perc_white <- models_nonlin_blk_wht$plots_white[[4]]

nh3_agric_perc_black <- models_nonlin_blk_wht$plots[[1]]
nox_commer_perc_black <- models_nonlin_blk_wht$plots[[2]]
nox_trans_perc_black <- models_nonlin_blk_wht$plots[[3]]
so2_energy_perc_black <- models_nonlin_blk_wht$plots[[4]]

```


## 2. Plots Economy variables

### 2a. Histograms

```{r}

dens_unempl <- data_master %>%
  ggplot(aes(x = perc_unemp)) +
  geom_density(aes(y =..scaled..)) +
  theme_minimal(base_size = 15) +
  ylab("") +
  xlab("% Unemployment") 

dens_poverty <- data_master %>%
  ggplot(aes(x = perc_pvert)) +
   geom_density(aes(y =..scaled..)) +
  theme_minimal(base_size = 15) +
  ylab("") +
  xlab("% Poverty")

dens_income <- data_master %>%
  ggplot(aes(x = fam_incom_adjs)) +
  geom_density(aes(y =..scaled..)) +
  theme_minimal(base_size = 15) +
  ylab("") +
  xlab("Median Family Income ($1000)") +
   scale_x_continuous(labels=function(x)x/1000, breaks = c(0, 25000, 50000, 75000, 100000, 125000, 150000))  

dens_propty <- data_master %>%
  ggplot(aes(x = propty_val_adjs)) +
   geom_density(aes(y =..scaled..)) +
  theme_minimal(base_size = 15) +
  ylab("") +
  xlab("Median Propty Value ($1000)") +
 scale_x_continuous(labels=function(x)x/1000)  
  
```


### 2b. Adjust axis 

Adjust the scale of the x-axis for income and property value

```{r}
# fix income axis scale

so2_indus_fam_incom_adjs <- so2_indus_fam_incom_adjs +
   scale_x_continuous(labels=function(x)x/1000, breaks = c(0, 25000, 50000, 75000, 100000, 125000, 150000))  

nox_energy_fam_incom_adjs <- nox_energy_fam_incom_adjs  +
   scale_x_continuous(labels=function(x)x/1000, breaks = c(0, 25000, 50000, 75000, 100000, 125000, 150000))

so2_energy_fam_incom_adjs <- so2_energy_fam_incom_adjs  +
   scale_x_continuous(labels=function(x)x/1000, breaks = c(0, 25000, 50000, 75000, 100000, 125000, 150000))

nh3_agric_fam_incom_adjs <- nh3_agric_fam_incom_adjs  +
   scale_x_continuous(labels=function(x)x/1000, breaks = c(0, 25000, 50000, 75000, 100000, 125000, 150000))

nox_trans_fam_incom_adjs <- nox_trans_fam_incom_adjs  +
   scale_x_continuous(labels=function(x)x/1000, breaks = c(0, 25000, 50000, 75000, 100000, 125000, 150000))

oc_resid_fam_incom_adjs <- oc_resid_fam_incom_adjs  +
   scale_x_continuous(labels=function(x)x/1000, breaks = c(0, 25000, 50000, 75000, 100000, 125000, 150000))

nox_commer_fam_incom_adjs <- nox_commer_fam_incom_adjs  +
   scale_x_continuous(labels=function(x)x/1000, breaks = c(0, 25000, 50000, 75000, 100000, 125000, 150000))

# fix property value axis scale

so2_indus_propty_val_adjs <- so2_indus_propty_val_adjs +
  scale_x_continuous(labels=function(x)x/1000) 

nox_energy_propty_val_adjs <- nox_energy_propty_val_adjs +
  scale_x_continuous(labels=function(x)x/1000) 

so2_energy_propty_val_adjs <- so2_energy_propty_val_adjs +
  scale_x_continuous(labels=function(x)x/1000) 

nh3_agric_propty_val_adjs <- nh3_agric_propty_val_adjs +
  scale_x_continuous(labels=function(x)x/1000) 

nox_trans_propty_val_adjs <- nox_trans_propty_val_adjs +
  scale_x_continuous(labels=function(x)x/1000) 

oc_resid_propty_val_adjs <- oc_resid_propty_val_adjs +
  scale_x_continuous(labels=function(x)x/1000)

nox_commer_propty_val_adjs <- nox_commer_propty_val_adjs +
  scale_x_continuous(labels=function(x)x/1000)
```


### 2b. Ensemble panel figure economic variables

```{r}

pdf(file = "curves_eco.pdf", width = 14, height = 20)            
 grid.arrange(dens_unempl,
              dens_poverty,
              dens_income,
              dens_propty,
   
              so2_indus_perc_unemp,
              so2_indus_perc_pvert,
              so2_indus_fam_incom_adjs,
              so2_indus_propty_val_adjs,
              
              so2_energy_perc_unemp,
              so2_energy_perc_pvert,
              so2_energy_fam_incom_adjs,
              so2_energy_propty_val_adjs,
              
              nox_energy_perc_unemp,
              nox_energy_perc_pvert,
              nox_energy_fam_incom_adjs,
              nox_energy_propty_val_adjs,
              
              nh3_agric_perc_unemp,
              nh3_agric_perc_pvert,
              nh3_agric_fam_incom_adjs,
              nh3_agric_propty_val_adjs,
              
              nox_trans_perc_unemp,
              nox_trans_perc_pvert,
              nox_trans_fam_incom_adjs,
              nox_trans_propty_val_adjs,
              
              oc_resid_perc_unemp,
              oc_resid_perc_pvert,
              oc_resid_fam_incom_adjs,
              oc_resid_propty_val_adjs,
              
              nox_commer_perc_unemp,
              nox_commer_perc_pvert,
              nox_commer_fam_incom_adjs,
              nox_commer_propty_val_adjs,
              
             nrow = 8, ncol = 4)
            # left = gridtext::richtext_grob("Industry: SO<sub>2</sub> (% chg)", rot=90,
             #gp=gpar(fontsize=18,font=8)))
dev.off() 

```

## 2. Plots Race variables

### 2a. Histograms

```{r}

dens_black <- data_master %>%
  ggplot(aes(x = perc_black)) +
  geom_density(aes(y =..scaled..)) +
  theme_minimal(base_size = 15) +
  ylab("") +
  xlab("% Black") +
  xlim(-3,86.5)

dens_white <- data_master %>%
  ggplot(aes(x = perc_white)) +
   geom_density(aes(y =..scaled..)) +
  theme_minimal(base_size = 15) +
  ylab("") +
  xlab("% White") +
  xlim(0, 104)

dens_asian <- data_master %>%
  ggplot(aes(x = perc_asian)) +
  geom_density(aes(y =..scaled..)) +
  theme_minimal(base_size = 15) +
  ylab("") +
  xlab("% Asian") +
  xlim(-0.28,45)


dens_am_nat <- data_master %>%
  ggplot(aes(x = perc_nativ)) +
   geom_density(aes(y =..scaled..)) +
  theme_minimal(base_size = 15) +
  ylab("") +
  xlab("% American Native") +
  xlim(-0.17, 95)

dens_hisp <- data_master %>%
  ggplot(aes(x = perc_hisp_org)) +
   geom_density(aes(y =..scaled..)) +
  theme_minimal(base_size = 15) +
  ylab("") +
  xlab("% Hispanic") +
  xlim(-0.65, 100)

```

### 2b. Edit axis of Hispanic
```{r}

so2_indus_perc_hisp_org <- so2_indus_perc_hisp_org +
   xlim(0,100) 

so2_energy_perc_hisp_org <- so2_energy_perc_hisp_org +
  xlim(0,100)

nox_energy_perc_hisp_org <- nox_energy_perc_hisp_org +
  xlim(0,100)

nh3_agric_perc_hisp_org  <- nh3_agric_perc_hisp_org  +
xlim(0,100)

nox_trans_perc_hisp_org <- nox_trans_perc_hisp_org +
xlim(0,100)

oc_resid_perc_hisp_org <-  oc_resid_perc_hisp_org +
xlim(0,100)

nox_commer_perc_hisp_org <- nox_commer_perc_hisp_org +
xlim(0,100)

```


### 2c. Ensemble panel figure race

```{r}

pdf(file = "curves_race.pdf", width = 14, height = 20)            
 grid.arrange(dens_white,
              dens_black,
              dens_asian,
              dens_am_nat,
              dens_hisp,
   
              so2_indus_perc_white,
              so2_indus_perc_black,
              so2_indus_perc_asian,
              so2_indus_perc_nativ,
              so2_indus_perc_hisp_org,
              
              so2_energy_perc_white,
              so2_energy_perc_black, 
              so2_energy_perc_asian,
              so2_energy_perc_nativ,
              so2_energy_perc_hisp_org,
              
              nox_energy_perc_white,
              nox_energy_perc_black,
              nox_energy_perc_asian,
              nox_energy_perc_nativ,
              nox_energy_perc_hisp_org,
              
              nh3_agric_perc_white,
              nh3_agric_perc_black, 
              nh3_agric_perc_asian,
              nh3_agric_perc_nativ,
              nh3_agric_perc_hisp_org,
              
              nox_trans_perc_white,
              nox_trans_perc_black, 
              nox_trans_perc_asian,
              nox_trans_perc_nativ,
              nox_trans_perc_hisp_org,
              
              oc_resid_perc_white,
              oc_resid_perc_black,
              oc_resid_perc_asian,
              oc_resid_perc_nativ,
              oc_resid_perc_hisp_org,
              
              nox_commer_perc_white,
              nox_commer_perc_black,  
              nox_commer_perc_asian,
              nox_commer_perc_nativ,
              nox_commer_perc_hisp_org,
              
             nrow = 8, ncol = 5)
            # left = gridtext::richtext_grob("Industry: SO<sub>2</sub> (% chg)", rot=90,
             #gp=gpar(fontsize=18,font=8)))
dev.off() 

```