---
title: "emissions maps"
author: "Yanelli Nunez"
date: "3/1/2022"
output:
 html_document:
    toc: TRUE
    toc_float: TRUE
    code_folding: hide
editor_options: 
  chunk_output_type: console
---

*Objective:* 

This code generates manuscript's Figure 1 (the emissions maps)

Note: you won't be able to run this code because the geometries are saved in my local drive

```{r include=FALSE}

rm(list=ls())

# 1a Declare root directory

project.folder <- paste0(print(here::here()),'/')

# add folder locations
source(paste0(project.folder,'0_00_create_folder_structure.R'))

# add file locations
source(paste0(file.locations.folder,'external_file_locations.R'))

# load packages
source(paste0(packages.folder,'packages_to_load.R'))

# load functions
 source(paste0(code.folder, 'models/functions.R'))

# load data
source(paste0(code.folder, 'data_loading.R'))

```


## 1. Load data

```{r}

class(data_full)
summary(data_full)

# geometries 
geometries <- readRDS(paste0(yanelli_external, "spatial_data/updated_generated_files/geometries_counties.rds")) %>%
  mutate(name = if_else(gisjoin == "G1200250", "Miami-Dade county", name), # rename Dade to Miami-Dade so it is consistent naming across the 4 years
         gisjoin = if_else(gisjoin == "G1200250", "G1200860", gisjoin),
         name = if_else(gisjoin == "G1200860", "Miami-Dade", name)) # rename Dade to Miami-Dade so it is consistent naming across the 4 years

class(geometries) 

data_sf <- geometries %>%
  dplyr::right_join(., data_full)

summary(data_sf)
class(data_sf)
```

## My color palette 

```{r}

my_cols_het <- c("#7fc97f", "#beaed4", "#fdc086", "#999999", "#386cb0", "#f0027f", "#bf5b17", "#377eb8", "#a65628")

my_purpls <- c("#efedf5", "#dadaeb", "#bcbddc", "#9e9ac8", "#807dba", "#6a51a3", "#54278f", "#3f007d", "#1D0633", "#000000")

my_greens <- c("#f7fcfd",
"#e5f5f9",
"#ccece6",
"#99d8c9",
"#66c2a4",
"#41ae76",
"#238b45",
"#006d2c",
"#00441b", "#000000")

my_blues <- c("#f7fbff",
"#deebf7",
"#c6dbef",
"#9ecae1",
"#6baed6",
"#4292c6",
"#2171b5",
"#08519c",
"#08306b", "#000000")

my_orang <- c(
"#fff5eb",
"#fee6ce",
"#fdd0a2",
"#fdae6b",
"#fd8d3c",
"#f16913",
"#d94801",
"#a63603",
"#7f2704",
"#000000") 

my_pinks <- c(
"#fff7f3",
"#fde0dd",
"#fcc5c0",
"#fa9fb5",
"#f768a1",
"#dd3497",
"#ae017e",
"#7a0177",
"#49006a",
"#000000"
)

my_grays <- c(
"#ffffff",
"#f7fbff",
"#f0f0f0",
"#d9d9d9",
#"ivory3",
"#bdbdbd",
"#969696",
"#737373",
"#525252",
"#252525",
"#000000"
)

my_browns <- c(
"cornsilk",
"wheat",
"burlywood",
"tan3",
#"peru",
"tan4",
"sienna",
"saddlebrown",
#"salmon4",
"#993404",
"#662506",
"#000000"  
)


```


## 3. Map

### 3a. Industry, SO2 emissions

```{r}

# function for mapping 
# mapping <- function(data, year){

start_time <- Sys.time()    
pdf(file = "so2_maps.pdf", width = 14, height = 6)

data_sf %>%
  filter(year %in% c(1970, 2010)) %>%
  mutate(decile = cut_number(so2_industry, 10)) %>%
  ggplot() +
  geom_sf(colour = alpha("antiquewhite3", 0.25), size = 0.7) +
  aes(fill = decile) +
  #scale_fill_viridis(name=SO[2]~"Industry", option = "viridis", discrete = TRUE)    
  scale_fill_manual(name=SO[2]~"Industry"~(kg/km^2/day), values = my_purpls, na.translate = FALSE) +
  facet_wrap( ~ year) +
  theme_void() +
   theme(legend.title=element_text(size=12))
  #       strip.text = element_text(size = 30)
  #      axis.line = element_blank(), 
  #      axis.text = element_blank(),
  #      axis.ticks = element_blank(), 
  #      axis.title = element_blank(),
  #      plot.title = element_text(size=20)) 

dev.off()
end_time <- Sys.time()
end_time - start_time  

#}


```

### 3a. Energy, SO2 emissions

```{r}


start_time <- Sys.time()    
pdf(file = "energy_so2_maps.pdf", width = 14, height = 6)

data_sf %>%
  filter(year %in% c(1970, 2010)) %>%
  mutate(decile = cut_number(so2_ener, 10)) %>%
  ggplot() +
  geom_sf(colour = alpha("antiquewhite3", 0.25), size = 0.7) +
  aes(fill = decile) +
  scale_fill_manual(name=SO[2]~"Energy"~(kg/km^2/day), values = my_orang, na.translate = FALSE) +
  facet_wrap( ~ year) +
  theme_void() +
   theme(legend.title=element_text(size=12))

dev.off()
end_time <- Sys.time()
end_time - start_time  

```


### 3a. Energy, NOx emissions

```{r}


start_time <- Sys.time()    
pdf(file = "energy_nox_maps.pdf", width = 14, height = 6)

data_sf %>%
  filter(year %in% c(1970, 2010)) %>%
  mutate(decile = cut_number(nox_ener, 10)) %>%
  ggplot() +
  geom_sf(colour = alpha("antiquewhite3", 0.25), size = 0.7) +
  aes(fill = decile) +
  scale_fill_manual(name=NO[x]~"Energy"~(kg/km^2/day), values = my_blues, na.translate = FALSE) +
  facet_wrap( ~ year) +
  theme_void() +
   theme(legend.title=element_text(size=12))

dev.off()
end_time <- Sys.time()
end_time - start_time  

```



### 3a. Agriculture, NH3 emissions

```{r}


start_time <- Sys.time()    
pdf(file = "agricul_nh3_maps.pdf", width = 14, height = 6)

data_sf %>%
  filter(year %in% c(1970, 2010)) %>%
  mutate(decile = cut_number(nh3_agric, 10)) %>%
  ggplot() +
  geom_sf(colour = alpha("antiquewhite3", 0.25), size = 0.7) +
  aes(fill = decile) +
  scale_fill_manual(name=NH[3]~"Agriculture"~(kg/km^2/day), values = my_greens, na.translate = FALSE) +
  facet_wrap( ~ year) +
  theme_void() +
   theme(legend.title=element_text(size=12))

dev.off()
end_time <- Sys.time()
end_time - start_time  

```


### 3a. Transportation, NOx emissions

```{r}

start_time <- Sys.time()    
pdf(file = "transport_nox_maps.pdf", width = 14, height = 6)

data_sf %>%
  filter(year %in% c(1970, 2010)) %>%
  mutate(decile = cut_number(nox_trans, 10)) %>%
  ggplot() +
  geom_sf(colour = alpha("antiquewhite3", 0.25), size = 0.7) +
  aes(fill = decile) +
  scale_fill_manual(name=NO[x]~"Transport"~(kg/km^2/day), values = my_grays, na.translate = FALSE) +
  facet_wrap( ~ year) +
  theme_void() +
   theme(legend.title=element_text(size=12))

dev.off()
end_time <- Sys.time()
end_time - start_time  

```


### 3a. Transportation, NOx emissions

```{r}

start_time <- Sys.time()    
pdf(file = "transport_nox_maps.pdf", width = 14, height = 6)

data_sf %>%
  filter(year %in% c(1970, 2010)) %>%
  mutate(decile = cut_number(nox_trans, 10)) %>%
  ggplot() +
  geom_sf(colour = alpha("antiquewhite3", 0.25), size = 0.7) +
  aes(fill = decile) +
  scale_fill_manual(name=NO[x]~"Transport"~(kg/km^2/day), values = my_grays, na.translate = FALSE) +
  facet_wrap( ~ year) +
  theme_void() +
   theme(legend.title=element_text(size=12))

dev.off()
end_time <- Sys.time()
end_time - start_time  

```


### 3a. Residencial, OC emissions

```{r}

start_time <- Sys.time()    
pdf(file = "residen_oc_maps.pdf", width = 14, height = 6)

data_sf %>%
  filter(year %in% c(1970, 2010)) %>%
  mutate(decile = cut_number(oc_residential, 10)) %>%
  ggplot() +
  geom_sf(colour = alpha("antiquewhite3", 0.25), size = 0.7) +
  aes(fill = decile) +
  scale_fill_manual(name="OC Residential"~(kg/km^2/day), values = my_pinks, na.translate = FALSE) +
  facet_wrap( ~ year) +
  theme_void() +
   theme(legend.title=element_text(size=12))

dev.off()
end_time <- Sys.time()
end_time - start_time  

```


### 3a. Commercial, NOx emissions

```{r}

start_time <- Sys.time()    
pdf(file = "commercial_nox_maps.pdf", width = 14, height = 6)

data_sf %>%
  filter(year %in% c(1970, 2010)) %>%
  mutate(decile = cut_number(nox_commercial, 10)) %>%
  ggplot() +
  geom_sf(colour = alpha("antiquewhite3", 0.25), size = 0.7) +
  aes(fill = decile) +
  scale_fill_manual(name=NO[x]~"Commercial"~(kg/km^2/day), values = my_browns, na.translate = FALSE) +
  facet_wrap( ~ year) +
  theme_void() +
   theme(legend.title=element_text(size=12))

dev.off()
end_time <- Sys.time()
end_time - start_time  

```
