---
title: "emissions maps"
author: "Yanelli Nunez"
date: "3/1/2022"
output:
 html_document:
    toc: TRUE
    toc_float: TRUE
    code_folding: hide
editor_options: 
  chunk_output_type: console
---

*Objective:* 

This code generates manuscript's Figure 1 (emissions difference from 1970 to 2010). This figure substituted the previous figures showing concentrations in the 1970 and 2010 time point (y_03_emissions_maps.Rmd). 


```{r include=FALSE}

rm(list=ls())

# 1a Declare root directory

project.folder <- paste0(print(here::here()),'/')

# add folder locations
source(paste0(project.folder,'0_00_create_folder_structure.R'))

# add file locations
source(paste0(file.locations.folder,'external_file_locations.R'))

# load packages
source(paste0(packages.folder,'packages_to_load.R'))

# load functions
 source(paste0(code.folder, 'models/functions.R'))

# load data
source(paste0(code.folder, 'data_loading.R'))

```


## 1. Load data & edit

```{r}

class(data_full)
summary(data_full)

diff_data <- data_full %>% 
   dplyr::select(gisjoin, year, state, county, name, epa_reg, tot_pop, 48:54) %>% 
  dplyr::filter(year %in% c("1970", "2010")) %>% 
    mutate(gisjoin = as.factor(gisjoin)) %>%
  group_by(state, gisjoin) %>%
   arrange(., year, .by_group = TRUE) %>%
    mutate(nox_trans_diff = (nox_trans - dplyr::lag(nox_trans)),
           nox_trans_diff = dplyr::lead(nox_trans_diff),
           nh3_agric_diff = (nh3_agric - dplyr::lag(nh3_agric)),
           nh3_agric_diff = dplyr::lead(nh3_agric_diff),
           so2_energy_diff = (so2_ener - dplyr::lag(so2_ener)),
           so2_energy_diff = dplyr::lead(so2_energy_diff),
           nox_energy_diff = (nox_ener - dplyr::lag(nox_ener)),
           nox_energy_diff = dplyr::lead(nox_energy_diff),
           so2_indus_diff = (so2_industry - dplyr::lag(so2_industry)),
           so2_indus_diff = dplyr::lead(so2_indus_diff),
           oc_resd_diff = (oc_residential - dplyr::lag(oc_residential)),
           oc_resd_diff = dplyr::lead(oc_resd_diff),
           nox_commer_diff = (nox_commercial - dplyr::lag(nox_commercial)),
           nox_commer_diff = dplyr::lead(nox_commer_diff)) %>% 
  ungroup() %>%
  mutate(nox_trans_relat_chg = nox_trans_diff/nox_trans*100, # relative change
         nh3_agric_relat_chg = nh3_agric_diff/nh3_agric*100,
         oc_resid_relat_chg = oc_resd_diff/oc_residential*100,
         nox_commer_relat_chg = nox_commer_diff/nox_commercial*100, 
         so2_energy_relat_chg = so2_energy_diff/so2_ener*100,
         nox_energy_relat_chg = nox_energy_diff/nox_ener*100, 
         so2_indus_relat_chg = so2_indus_diff/so2_industry*100) %>%
   as.data.frame() %>%
  dplyr::filter(year == "1970") %>% 
  dplyr::select(1:7, 15:28) %>%
  mutate_at("so2_indus_relat_chg",  ~replace(., is.nan(.), 0)) %>% 
  mutate(energy_so2_breaks = 
                        if_else(so2_energy_relat_chg >= -100 & so2_energy_relat_chg <= -80, "[-100, -80]", 
                        if_else(so2_energy_relat_chg > -80 & so2_energy_relat_chg <= -60, "(-80, -60]", 
                        if_else(so2_energy_relat_chg > -60 & so2_energy_relat_chg <= 0, "(-60, 0]", 
                        if_else(so2_energy_relat_chg > 0 & so2_energy_relat_chg <= 300, "(0, 300]", "> 300"))))) %>% 
  mutate(energy_so2_breaks = factor(energy_so2_breaks, levels = c("[-100, -80]", "(-80, -60]", "(-60, 0]", 
                                                                  "(0, 300]", "> 300"))) %>%
   mutate(energy_nox_breaks = 
                        if_else(nox_energy_relat_chg >= -100 & nox_energy_relat_chg <= -80, "[-100, -75]", 
                        if_else(nox_energy_relat_chg > -80 & nox_energy_relat_chg <= -60, "(-75, -50]", 
                        if_else(nox_energy_relat_chg > -60 & nox_energy_relat_chg <= 0, "(-50, 0]", 
                        if_else(nox_energy_relat_chg > 0 & nox_energy_relat_chg <= 300, "(0, 300]", "> 300"))))) %>% 
  mutate(energy_nox_breaks = factor(energy_nox_breaks, levels = c("[-100, -75]", "(-75, -50]", "(-50, 0]", 
                                                                  "(0, 300]", "> 300"))) %>% 
  mutate(oc_resid_breaks = 
                        if_else(oc_resid_relat_chg >= -25 & oc_resid_relat_chg <= 0, "[-25, 0]", 
                        if_else(oc_resid_relat_chg > 0 & oc_resid_relat_chg <= 10, "(0, 10]", 
                        if_else(oc_resid_relat_chg > 10 & oc_resid_relat_chg <= 20, "(10, 20]", 
                        if_else(oc_resid_relat_chg > 20 & oc_resid_relat_chg <= 100, "(20, 100]", "> 100"))))) %>% 
  mutate(oc_resid_breaks = factor(oc_resid_breaks, levels = c("[-25, 0]", "(0, 10]", "(10, 20]", 
                                                                  "(20, 100]", "> 100"))) %>% 
  mutate(nox_commer_breaks = 
                        if_else(nox_commer_relat_chg >= -77 & nox_commer_relat_chg <= -50, "[-77, -50]", 
                        if_else(nox_commer_relat_chg > -50 & nox_commer_relat_chg <= -25, "(-50, -25]", 
                        if_else(nox_commer_relat_chg > -25 & nox_commer_relat_chg <= 0, "(-25, 0]", 
                        if_else(nox_commer_relat_chg > 0 & nox_commer_relat_chg <= 50, "(0, 50]", "> 50"))))) %>% 
  mutate(nox_commer_breaks = factor(nox_commer_breaks, levels = c("[-77, -50]", "(-50, -25]", "(-25, 0]", 
                                                                  "(0, 50]", "> 50"))) %>% 
  mutate(so2_ind_breaks = 
                        if_else(so2_indus_relat_chg >= -100 & so2_indus_relat_chg <= -90, "[-100, -90]", 
                        if_else(so2_indus_relat_chg > -90 & so2_indus_relat_chg <= -80, "(-90, -80]", 
                        if_else(so2_indus_relat_chg > -80 & so2_indus_relat_chg <= -50, "(-80, -50]", 
                        if_else(so2_indus_relat_chg > -50 & so2_indus_relat_chg <= -20, "(-50, -20]", "(-20, 0]"))))) %>% 
  mutate(so2_ind_breaks = factor(so2_ind_breaks, levels = c("[-100, -90]", "(-90, -80]", "(-80, -50]", 
                                                                  "(-50, -20]", "(-20, 0]"))) %>% 
   mutate(nh3_agr_breaks = 
                        if_else(nh3_agric_relat_chg >= 1 & nh3_agric_relat_chg <= 25, "[1, 25]", 
                        if_else(nh3_agric_relat_chg > 25 & nh3_agric_relat_chg <= 50, "(25, 50]", 
                        if_else(nh3_agric_relat_chg > 50 & nh3_agric_relat_chg <= 75, "(50, 75]", 
                        if_else(nh3_agric_relat_chg > 75 & nh3_agric_relat_chg <= 100, "(75, 100]", "(100, 185]"))))) %>% 
  mutate(nh3_agr_breaks = factor(nh3_agr_breaks, levels = c("[1, 25]", "(25, 50]", "(50, 75]", 
                                                                  "(75, 100]", "(100, 185]"))) %>% 
  mutate(nox_trans_breaks = 
                        if_else(nox_trans_relat_chg >= -80 & nox_trans_relat_chg <= -70, "[-80, -70]", 
                        if_else(nox_trans_relat_chg > -70 & nox_trans_relat_chg <= -60, "(-70, -60]", 
                        if_else(nox_trans_relat_chg > -60 & nox_trans_relat_chg <= -50, "(-60, -50]",
                        if_else(nox_trans_relat_chg > -50 & nox_trans_relat_chg <= -1, "(-50, -1]", "(-1, 3]"))))) %>% 
  mutate(nox_trans_breaks = factor(nox_trans_breaks, levels = c("[-80, -70]", "(-70, -60]", "(-60, -50]", 
                                                                  "(-50, -1]", "(-1, 3]")))
  
summary(diff_data)
  

```

## 2. Spatial Geometries

Note that the spatial geometries are not included in the GitHub repository because of size limits. You can download the county geometries directly for years 1970 and 2010 directly from nhgis.org.

```{r}

# geometries 
geometries <- readRDS(paste0(yanelli_external, "spatial_data/updated_generated_files/geometries_counties.rds")) %>%
  mutate(name = if_else(gisjoin == "G1200250", "Miami-Dade county", name), # rename Dade to Miami-Dade so it is consistent naming across the 4 years
         gisjoin = if_else(gisjoin == "G1200250", "G1200860", gisjoin),
         name = if_else(gisjoin == "G1200860", "Miami-Dade", name)) # rename Dade to Miami-Dade so it is consistent naming across the 4 years

class(geometries) 
geometries$geometry


data_sf <- geometries %>%
  dplyr::right_join(., diff_data)

summary(data_sf)
class(data_sf)

```


## 3. My color palette 

```{r}

my_cols_het <- c("#7fc97f", "#beaed4", "#fdc086", "#999999", "#386cb0", "#f0027f", "#bf5b17", "#377eb8", "#a65628")

my_purpls <- c("#efedf5", 
#"#dadaeb", 
"#bcbddc", 
#"#9e9ac8", 
"#807dba", 
#"#6a51a3", 
#"#54278f", 
"#3f007d", 
#"#1D0633", 
"#000000")

my_greens <- c( "#f7fcfd",
  "#99d8c9", 
  "#238b45",
  "#00441b", 
  "#000000")

my_blues <- c( "#f7fbff",
# "#deebf7",
# "#c6dbef",
 "#9ecae1",
# "#6baed6",
 "#4292c6",
# "#2171b5",
 "#08519c",
 #"#08306b", 
 "#000000")

my_orang <- c(
  "#fff5eb",
# "#fee6ce",
#  "#fdd0a2",
  "#fdae6b",
#  "#fd8d3c",
  "#f16913",
#  "#d94801",
  "#a63603",
#  "#7f2704",
  "#000000") 

my_pinks <- c(
 "#fff7f3",
#  "#fde0dd",
#  "#fcc5c0",
  "#fa9fb5",
#  "#f768a1",
#  "#dd3497",
  "#ae017e",
 # "#7a0177",
  "#49006a",
  "#000000"
)

my_grays <- c(
  "#ffffff",
#  "#f7fbff",
#  "#f0f0f0",
#  "#d9d9d9",
  #"ivory3",
  "#bdbdbd",
 # "#969696",
  "#737373",
  "#525252",
 # "#252525",
  "#000000"
)

my_browns <- c(
"cornsilk",
#  "wheat",
  "burlywood",
 # "tan3",
  #"peru",
  "tan4",
#  "sienna",
  #"saddlebrown",
   "salmon4",
#  "#993404",
  "#662506",
  "#000000" 
)


```


## 3. Map

### 3a. Industry, SO2 emissions

```{r}

# function for mapping 
# mapping <- function(data, year){

start_time <- Sys.time()    
pdf(file = "so2_maps.pdf", width = 14, height = 6)

data_sf %>%
 # filter(state == "New York") %>%
  ggplot() +
  geom_sf(colour = alpha("black", 0.25), size = 0.7) +
  aes(fill = so2_ind_breaks) +
  #scale_fill_viridis(name=SO[2]~"Industry", option = "viridis", discrete = TRUE)    
  scale_fill_manual(name=SO[2]~"Industry", values = my_grays, na.translate = FALSE) +
  theme_void() +
   theme(legend.title=element_text(size=12))
  #       strip.text = element_text(size = 30)
  #      axis.line = element_blank(), 
  #      axis.text = element_blank(),
  #      axis.ticks = element_blank(), 
  #      axis.title = element_blank(),
  #      plot.title = element_text(size=20)) 

dev.off()
end_time <- Sys.time()
end_time - start_time  

#}


```


### 3a. Energy, SO2 emissions


```{r}


start_time <- Sys.time()    
pdf(file = "energy_so2_maps.pdf", width = 14, height = 6)

data_sf %>%
 # mutate(decile = cut_interval(so2_energy_relat_chg, 5)) %>%
  ggplot() +
  geom_sf(colour = alpha("black", 0.25), size = 0.7) +
  aes(fill = energy_so2_breaks) +
  scale_fill_manual(name=SO[2]~"Energy", values = my_blues, na.translate = FALSE) +
  theme_void() +
   theme(legend.title=element_text(size=12))

dev.off()
end_time <- Sys.time()
end_time - start_time  

```


### 3a. Energy, NOx emissions


```{r}


start_time <- Sys.time()    
pdf(file = "energy_nox_maps.pdf", width = 14, height = 6)

data_sf %>%
  filter(year %in% c(1970, 2010)) %>%
  ggplot() +
  geom_sf(colour = alpha("black", 0.25), size = 0.7) +
  aes(fill = energy_nox_breaks) +
  scale_fill_manual(name=NO[x]~"Energy", values = my_pinks, na.translate = FALSE) +
  theme_void() +
   theme(legend.title=element_text(size=12))

dev.off()
end_time <- Sys.time()
end_time - start_time  

```


### 3a. Agriculture, NH3 emissions


```{r}


start_time <- Sys.time()    
pdf(file = "agricul_nh3_maps.pdf", width = 14, height = 6)

data_sf %>%
  ggplot() +
  geom_sf(colour = alpha("black", 0.25), size = 0.7) +
  aes(fill = nh3_agr_breaks) +
  scale_fill_manual(name=NH[3]~"Agriculture", values = my_greens, na.translate = FALSE) +
  theme_void() +
   theme(legend.title=element_text(size=12))

dev.off()
end_time <- Sys.time()
end_time - start_time  

```


### 3a. Transportation, NOx emissions

```{r}

start_time <- Sys.time()    
pdf(file = "transport_nox_maps.pdf", width = 14, height = 6)

data_sf %>%
  ggplot() +
  geom_sf(colour = alpha("black", 0.25), size = 0.7) +
  aes(fill = nox_trans_breaks) +
  scale_fill_manual(name=NO[x]~"Transport", values = my_orang, na.translate = FALSE) +
  theme_void() +
   theme(legend.title=element_text(size=12))

dev.off()
end_time <- Sys.time()
end_time - start_time  


```


### 3a. Residencial, OC emissions


```{r}

start_time <- Sys.time()    
pdf(file = "residen_oc_maps.pdf", width = 14, height = 6)

data_sf %>%
  ggplot() +
  geom_sf(colour = alpha("black", 0.25), size = 0.7) +
  aes(fill = oc_resid_breaks) +
  scale_fill_manual(name="OC Residential", values = my_browns, na.translate = FALSE) +
  theme_void() +
   theme(legend.title=element_text(size=12))

dev.off()
end_time <- Sys.time()
end_time - start_time  

```


### 3a. Commercial, NOx emissions

**need to make breaks manually**

```{r}

start_time <- Sys.time()    
pdf(file = "commercial_nox_maps.pdf", width = 14, height = 6)

data_sf %>%
  ggplot() +
  geom_sf(colour = alpha("black", 0.25), size = 0.7) +
  aes(fill = nox_commer_breaks) +
  scale_fill_manual(name=NO[x]~"Commercial", values = my_purpls, na.translate = FALSE) +
  theme_void() +
   theme(legend.title=element_text(size=12))

dev.off()
end_time <- Sys.time()
end_time - start_time  

```


### 4. Mapping using Leaflet

```{r}

crs_wgs84 <- "+proj=longlat +datum=WGS84 +no_defs +ellps=WGS84 +towgs84=0,0,0"

# prepare data and select single year
data_nh3 <- data_sf %>% 
  dplyr::filter(year == 2010) %>%
  mutate(nh3_cat = if_else(nh3_agric <= 0.5, "< 0.5",
                   if_else(nh3_agric > 0.5 & nh3_agric <= 1, "(0.5 - 1]",
                   if_else(nh3_agric > 1 & nh3_agric <=2, "(1 - 2]", 
                   if_else(nh3_agric > 2 & nh3_agric <= 5, "(2 - 5]", "> 5"))))) %>%
  mutate(nh3_cat = factor(nh3_cat, levels = c("< 0.5", "(0.5 - 1]", "(1 - 2]","(2 - 5]", "> 5"))) %>% 
  as.data.frame() %>% 
       st_as_sf() %>%
  st_collection_extract(.) %>% # to convert the geometries from geometry collection back to multipolygon. The polygons were converted to polygons during the process of aggregating the years. In leaflet, we can't plot geometry collections
  st_transform(crs_wgs84)

class(data_nh3)
data_nh3$geometry

# Specify color pallet
color_pal <- colorFactor(palette = c("#f7fcfd",
#"#e5f5f9",
#"#ccece6",
"#99d8c9",
#"#66c2a4",
#"#41ae76",
"#238b45",
#"#006d2c",
"#00441b", 
"#000000"),
domain = data_nh3$nh3_cat)

# Map
leaflet(data = data_nh3) %>%
        addTiles() %>%
        setView(lng = -99.01, lat = 41.83, zoom = 4.5) %>%
        addPolygons(data = data_nh3, 
                   color = "black",
                  weight = 0.3,
               fillColor = ~color_pal(data_nh3$nh3_cat),
             fillOpacity = 0.7,
                  # popup = polyPopup,
  highlight = highlightOptions(weight = 2,
                               color = "blue",
                               bringToFront = TRUE)) %>%
                                addLegend(pal = color_pal,
                                          values = data_nh3$nh3_cat, 
                                          opacity = 0.9, 
                                          title = "kg/km<sup>2</sup>/day", 
                                          position = "topright")
```

