---
title: "j_03_emissions_per_county_data_prep"
author: "Jaime Benavides"
date: "3/3/2021"
output: html_document
editor_options: 
  chunk_output_type: console
---


```{r include=FALSE}

rm(list=ls())

# 1a Declare root directory

project.folder <- paste0(print(here::here()),'/')

# add folder locations
source(paste0(project.folder,'0_00_create_folder_structure.R'))

# add file locations
#source(paste0(file.locations.folder,'file_locations.R'))

# load packages
source(paste0(packages.folder,'packages_to_load.R'))
```

## 1. Load geographical context data (County shapefiles and bounding box of continental US)  

```{r context, echo=FALSE}
# create bounding box for continental US
us_cont_bbox <- sf::st_as_sfc(sf::st_bbox(c(xmin = -130, xmax = -60, ymax = 51, ymin = 21), crs = 4326)) 
us_counties_over_time_sf <- readRDS(paste0(data.folder, "generated_data/", "us_counties_over_time_sf.rds")) %>%
  sf::st_transform(4326)
```

## 2. Loading pollutant/sector emissions data. This dataset contains the emissions for each year of interest (1970 to 2017) for each centroid (points) within the CEDS grid. There is a previous step to produce this dataset where data is extracted from the global coverage NETCDF and converted to the units of analysis (kg km-2 day-1). That step takes place at script j_02_get_emis_sector_pollutant.R. Below we do an example with SO2 and energy. 

```{r emis, echo=FALSE}
# choose pollutant  and sector name
# in this study you need to generate all the following combinations 
# pollutant: "NOx" sector_name: ener (Energy)
# pollutant: "NOx" sector_name: tr (On-road transport)
# pollutant: "NOx" sector_name: commercial (Commercial combustion)
# pollutant: "SO2" sector_name: ener (Energy)
# pollutant: "SO2" sector_name: industry (Industry)
# pollutant: "NH3" sector_name:  agric (Agriculture)
# pollutant: "OC"  sector_name: residential (Residential Combustion)
# example with S02 and energy
pollutant <- "SO2"
sector_name <- "ener"
emissions_data_path <- paste0(data.folder, "emissions_data/", "emis_", tolower(pollutant), "_",  sector_name, "_centroids.rds")
emis_pollutant_sector_centroids <- readRDS(emissions_data_path) %>%
  sf::st_transform(4326) %>%
  sf::st_intersection(us_cont_bbox) 
```

## 3. Create a regular grid given the centroids of the CEDS emissions data. The CEDS grid is 0.5 degrees times 0.5 degrees. We build this grid here in order to have grid cells (polygons shapes) for the estimation of emissions per county weighted by area   

```{r grid, echo=FALSE}
# create grid
cellsize <- 0.5 # CEDS grid is 0.5 degrees
emis_grid <- sf::st_sf(geom=sf::st_make_grid(
  sf::st_as_sfc(
    sf::st_bbox(emis_pollutant_sector_centroids) + 
      c(-cellsize/2, -cellsize/2,
        cellsize/2, cellsize/2)),
  what="polygons", cellsize=cellsize), crs=4326) 

# join spatially with centroids data (including emissions) 
emis <- sf::st_join(emis_grid, emis_pollutant_sector_centroids) %>%
na.omit()
```

## 4. Estimate the emissions per county by using the average weighted by area method contained in the package areal for each year of interest   

```{r areal_mean, echo=FALSE}
years <- c("1970", "1980", "1990", "2000", "2010", "2017")
for(year_eval in years){

# subset emissions  
emis_year <- emis[, c("i", "j", "lat", "lon", "within_us", "geometry", "id", year_eval)] 

# transform counties and emissions to planar CRS
target <- us_counties_over_time_sf %>%
  dplyr::filter(Year == year_eval) %>%
  sf::st_transform(2163)

source <- emis_year[,c("id", year_eval)] %>%
  sf::st_transform(2163) %>%
  dplyr::rename(pollutant_sector = year_eval)

# estimate emissions averaged by area
assign(paste0("pollutant_sector_", year_eval),areal::aw_interpolate(sf::st_make_valid(target), tid = GISJOIN, source = source, sid = id,
                                    weight = "sum", output = "sf", intensive = "pollutant_sector"))
}
# put together results for each year
emis_county <- rbind(pollutant_sector_1970, pollutant_sector_1980, pollutant_sector_1990, pollutant_sector_2000, pollutant_sector_2010, pollutant_sector_2017)
```

## 5. Save data  

```{r emis, echo=FALSE}
emis_county_df <- emis_county
sf::st_geometry(emis_county_df) <- NULL
saveRDS(emis_county_df, paste0(data.folder, "emissions_data/", "emis_", tolower(pollutant), "_",  sector_name,"_county_df.rds"))
```

