---
title: "Urbanicity_data"
author: "Yanelli Nunez"
date: "6/7/2021"
output:
 html_document:
    toc: TRUE
    toc_float: TRUE
    code_folding: hide
editor_options: 
  chunk_output_type: console
---

Objective:

To create an urbanicity variable based on population
To create a climate regions variable classifying states into the NOAA climate regions
To create a EPA regions variable classifying states into the corresponding EPA region


```{r include=FALSE}

rm(list=ls())

# 1a Declare root directory

project.folder <- paste0(print(here::here()),'/')

# add folder locations
source(paste0(project.folder,'0_00_create_folder_structure.R'))

# add file locations
source(paste0(file.locations.folder,'external_file_locations.R'))

# load packages
source(paste0(packages.folder,'packages_to_load.R'))


```

## 1. Load data

```{r}

# data
data <- read_csv(paste0(data.folder, "generated_data/data_indp_varib_2.csv")) 

# spatial files for mapping 
geometries <- readRDS(paste0(yanelli_external, "spatial_data/updated_generated_files/geometries_counties.rds")) %>%
  clean_names() 

```


## 2. Urbanicity classification

- The classification from National Center for Health Statistics for Urbanicity is fixed for 2013 but since counties have been changing and evolving, here I am using a classification based on population instead. 

county pop ≥ 50,000 people --> metropolitan
10,000 ≤ county pop < 50,000 --> micropolitan
county pop < 10,000 non-urban

I obtain the classification from the link below, page 3

https://www.cdc.gov/nchs/data/series/sr_02/sr02_166.pdf


```{r}

data_urbanicity <- data  %>%
  mutate(urbanicity = as.factor(if_else(tot_pop >= 50000, "metropolitan",
                                        if_else(tot_pop < 50000 & tot_pop >= 10000, "micropolitan",
                                                if_else(tot_pop < 10000, "non-urban", "metropolitan")))))

summary(data_urbanicity)

```

## 3. Climate regions

climate regions https://www.ncdc.noaa.gov/monitoring-references/maps/us-climate-regions.php

```{r}

data_climate <- data_urbanicity %>%
  mutate(climate_region = as.factor(case_when(state %in% c("Idaho","Oregon","Washington") ~ "Northwest",
                                    state %in% c("Arkansas","Kansas","Louisiana","Mississippi","Oklahoma","Texas") ~ "South",
                                    state %in% c("Alabama","Florida","Georgia", "North Carolina", "South Carolina", "Virginia") ~ "Southeast",
                                    state %in% c("Connecticut", "District Of Columbia", "Delaware", "Maine", "Maryland", "Massachusetts", "New Hampshire", "New Jersey", "New York", "Pennsylvania", "Rhode Island", "Vermont") ~ "Northeast",
                                    state %in% c("Iowa", "Michigan", "Minnesota", "Wisconsin") ~ "East_North_Central",
                                    state %in% c("Illinois", "Indiana", "Kentucky", "Missouri", "Ohio", "Tennessee", "West Virginia") ~ "Central",
                                    state %in% c("Arizona", "Colorado", "New Mexico", "Utah") ~ "Southwest",
                                    state %in% c("California", "Nevada") ~ "West",
                                    state %in% c("Montana", "Nebraska", "North Dakota", "South Dakota", "Wyoming") ~ "West_North_Central")))
                                    
summary(data_climate)
unique(data_climate$climate_region)

```

## 4. EPA regions

```{r}

epa_region <- data_climate %>%
  mutate(epa_reg = case_when(climate_region == "Northwest"        ~ "region_10",
                             climate_region == "West" | state  ==  "Arizona" ~ "region_9",
                             climate_region == "West_North_Central" & state != "Nebraska" | state %in% c("Utah", "Colorado") ~ "region_8",
                             state %in% c("Nebraska", "Kansas", "Iowa", "Missouri") ~ "region_7",
                             state %in% c("New Mexico", "Texas", "Oklahoma", "Arkansas", "Louisiana") ~ "region_6",
                             climate_region == "East_North_Central" & state != "Iowa" | state %in% c("Illinois", "Ohio", "Indiana") ~ "region_5",
                             climate_region == "Southeast" & state != "Virginia" | state %in% c("Mississippi", "Tennessee", "Kentucky")  ~ "region_4",
                             climate_region == "Northeast" | state %in% c("Virginia", "West Virginia") ~ "region_1_3")) %>%
  mutate(year = as.factor(year), 
      epa_reg = factor(epa_reg, levels = c("region_1_3", "region_4", "region_5", "region_6", "region_7", "region_8", "region_9", "region_10")))


summary(epa_region)

unique(epa_region$epa_reg)
levels(epa_region$epa_reg)
```

## 5. Save data file 

```{r}

saveRDS(epa_region, paste0(data.folder, "generated_data/data_indp_varib_3.rds")) 

```

## 6. Visualization

```{r}

data_sf <- geometries %>%
  dplyr::right_join(., epa_region, by = c("year", "state", "gisjoin"))

summary(data_sf)
class(data_sf)
```

### 6a. My color palette 

```{r}

my_purpls <- c("#efedf5", "#dadaeb", "#bcbddc", "#9e9ac8", "#807dba", "#6a51a3", "#54278f", "#3f007d", "#1D0633", "#000000")

```

### 6b. Urbanicity maps

Urbanicity maps

This takes a bit to run. It generates the two urbanicity maps that are part of the manuscript supplement materials

```{r}

library(viridisLite)

# 2000
start_time <- Sys.time()    

pdf(file = "Urbanicity_2000.pdf", width = 11, height = 8.5)
data_sf %>%
  filter(year == 2000) %>%
  ggplot() +
  geom_sf(colour = alpha("antiquewhite3", 1/2), size = 0.2) +
  aes(fill = urbanicity) +
  #scale_fill_viridis(name="Urbanicity", option = "magma")  +
  scale_fill_manual(name = "Urbancity", values = c("chocolate4", "darkgoldenrod3", "chartreuse4")) +
  theme_void() +
  ggtitle("2000") +
  theme(axis.line = element_blank(), 
        axis.text = element_blank(),
        axis.ticks = element_blank(), 
        axis.title = element_blank(),
        plot.title = element_text(size=20)) 
dev.off()

end_time <- Sys.time()
end_time - start_time  

# 1970
start_time <- Sys.time()    

pdf(file = "Urbanicity_1970.pdf", width = 11, height = 8.5)
data_sf %>%
  filter(year == 1970) %>%
  ggplot() +
  geom_sf(colour = alpha("antiquewhite3", 1/2), size = 0.2) +
  aes(fill = urbanicity) +
  #scale_fill_viridis(name="Urbanicity", option = "magma")  +
  scale_fill_manual(name = "Urbancity", values = c("chocolate4", "darkgoldenrod3", "chartreuse4")) +
  theme_void() +
  ggtitle("1970") +
  theme(axis.line = element_blank(), 
        axis.text = element_blank(),
        axis.ticks = element_blank(), 
        axis.title = element_blank(),
        plot.title = element_text(size=20)) 
dev.off()

end_time <- Sys.time()
end_time - start_time  
```


### 6c. EPA regions map

EPA regions are consisted throughout years. So here I plot 2000 to visualize

```{r}

my_cols <- c("#7fc97f", "#beaed4", "#fdc086", "#999999", "#386cb0", "#f0027f", "#bf5b17", "#377eb8", "#a65628")

pdf(file = "epa_regions.pdf", width = 11, height = 8.5)
data_sf %>%
  filter(year == 2010) %>%
  ggplot() +
  geom_sf(colour = alpha("antiquewhite3", 1/2), size = 0.2) +
  aes(fill = epa_reg) +
  scale_fill_manual(name = "EPA regions", values = my_cols) +
  theme_void() +
  theme(axis.line = element_blank(), 
        axis.text = element_blank(),
        axis.ticks = element_blank(), 
        axis.title = element_blank(),
        plot.title = element_text(size=20)) 
dev.off()

```

