---
title: "Inflation adjustments"
author: "Yanelli Nunez"
date: "3/25/2021"
output:
 html_document:
    toc: TRUE
    toc_float: TRUE
    code_folding: hide
editor_options: 
  chunk_output_type: console
---
Objective:

To adjust all dollar variables to 2010 dollars


```{r include=FALSE}

rm(list=ls())

# 1a Declare root directory

project.folder <- paste0(print(here::here()),'/')

# add folder locations
source(paste0(project.folder,'0_00_create_folder_structure.R'))

# add file locations
#source(paste0(file.locations.folder,'file_locations.R'))

# load packages
source(paste0(packages.folder,'packages_to_load.R'))


```

## 1. Load data

```{r}

# master dataset 
data <- readRDS(paste0(data.folder, "generated_data/data_indp_varib_3.rds"))

summary(data)

# inflation adjustments factors 

cpi <- readxl::read_excel(paste0(data.folder, "census_data/CPI_U_RS_edit.xlsx")) %>%
  clean_names() %>%
  dplyr::filter(year %in% c(1969, 1979, 1989, 1999, 2010, 2012)) # Remember census data are for the year prior (e.g. 2000 census has data from 1999)

```

## 2. Calculate adjusted estimates

Note: All dollar variables will be inflation adjusted to 2010 using the cpi-u-rs index. Using this index the formulate to adjust is as follow:

Example for a 1990 adjustment to 2010 dollars

inflation-adjusted estimate = 1990 estimate*(2010 cpi-u-rs/1990 cpi-u-rs)

Form more details please see the link below: 
https://www.census.gov/topics/income-poverty/income/guidance/current-vs-constant-dollars.html

### a. Estimate index fraction
-All dollar variables will be adjusted to 2010 dollars
-The index fraction is: 2010 cpi-u-rs/cpi-u-rs of the year to be adjusted
-2010 cpi-u-rs is **320.40**
```{r}

cpi_index <- cpi %>%
  mutate(index_fraction = 320.40/cpi_u_rs_index)

knitr::kable(cpi_index)

```

## b. Adjust to 2017 dollars
-Numbers use for the adjustment below are from the index_fraction in the cpi knitr table above
-Remember census data is for the year prior (e.g. 2000 census has data from 1999)
-Data for 2008-2012 ACS (label as 2010 in this data) is adjusted to 2012. So here I'm adjusting the 2012 dollars to 2010 dollars
```{r}

data_adjs <- data %>%
  group_by(year) %>%
  mutate(fam_incom_adjs = if_else(year == "1970", fam_income*5.2610837,
                          if_else(year == "1980", fam_income*2.8031496,
                          if_else(year == "1990", fam_income*1.6988335,
                          if_else(year == "2000", fam_income*1.3098937,
                          if_else(year == "2010", fam_income*0.9493333, 0))))),
         propty_val_adjs = if_else(year == "1970", propty_val*5.2610837,
                           if_else(year == "1980", propty_val*2.8031496,
                           if_else(year == "1990", propty_val*1.6988335,
                           if_else(year == "2000", propty_val*1.3098937,
                           if_else(year == "2010", propty_val*0.9493333, 0))))))

summary(data_adjs)

```

### c. Add variables to master data

```{r}

saveRDS(data_adjs, paste0(data.folder, "generated_data/data_indp_varib_3_inflat_adjst.csv"))

```

