---
title: "y_01_county_data_prep"
author: "Yanelli Nunez"
date: "2/23/2021"
output: html_document
editor_options: 
  chunk_output_type: console
---

Objective: 

- To estimate population density for each county. I first estimate the county area then used that and population to estimate county population density


```{r include=FALSE}

rm(list=ls())

# 1a Declare root directory

project.folder <- paste0(print(here::here()),'/')

# add folder locations
source(paste0(project.folder,'0_00_create_folder_structure.R'))

# add file locations
source(paste0(file.locations.folder,'external_file_locations.R'))

# load packages
source(paste0(packages.folder,'packages_to_load.R'))
```

## 1. Define the projection for the spatial data

```{r crs, include=FALSE}

crs_wgs84 <- "+proj=longlat +datum=WGS84 +no_defs +ellps=WGS84 +towgs84=0,0,0"

```

## 2. Spatial data files 

Spatial data files for the US and counties. I obtained this files from: 

https://www.census.gov/geographies/mapping-files/time-series/geo/carto-boundary-file.html


```{r pressure, echo=FALSE}
# create bounding box for continental US
us_cont_bbox <- sf::st_as_sfc(sf::st_bbox(c(xmin = -130, xmax = -60, ymax = 51, ymin = 21), crs = sf::st_crs(crs_wgs84))) %>%
  sf::st_transform(crs_wgs84)
  
# US continuous boundaries which I used in the following code chunck
us_cont_bds <- sf::st_read(paste0(yanelli_external, "spatial_data/cb_2018_us_nation_20m/cb_2018_us_nation_20m.shp")) %>%
  sf::st_transform(crs_wgs84) %>%
  sf::st_intersection(us_cont_bbox) %>%
  dplyr::select(geometry)

# US states which I used in the following code chunk to add a "States" column to the 2010 spatial dataset 
us_states <- sf::st_read(paste0(yanelli_external, "spatial_data/cb_2018_us_state_20m/cb_2018_us_state_20m.shp")) %>%
  sf::st_transform(crs_wgs84) %>%
  dplyr::rename(State = NAME)
  sf::st_intersection(us_cont_bds) %>%
  dplyr::select(State)

data <- readRDS(paste0(data.folder, "generated_data/data_indp_varib_3_inflat_adjst.rds")) 


```


## 3. Estimate county area
- Gather spatial attributes for counties for each year
- Estimate county area

Note: this data is saved in my local drive. I download this data from the nhgis website


```{r counties_over_time, echo=FALSE}

sf_use_s2(FALSE)  # the sf package was updated since Jaime created this code. This function specifies to use the older form of sf. Otherwise the code doesn't work

us_counties_1970 <- sf::st_read(paste0(yanelli_external, "spatial_data/nhgis0018_shape/nhgis0018_shapefile_tl2000_us_county_1970/US_county_1970.shp")) %>%
  sf::st_transform(crs_wgs84) %>%
  sf::st_make_valid() %>% # to 'clean' polygons. See https://r-spatial.org/r/2017/03/19/invalid.html
  sf::st_intersection(us_cont_bds) %>% # US continental boundaries
  dplyr::rename(State = STATENAM) %>%
  dplyr::rename(Name = NHGISNAM) %>%
  dplyr::rename(Year = DECADE) %>%
  dplyr::select(Name, GISJOIN, State, Year) 

us_counties_1980 <- sf::st_read(paste0(yanelli_external, "spatial_data/nhgis0018_shape/nhgis0018_shapefile_tl2000_us_county_1980/US_county_1980.shp")) %>%
  sf::st_transform(crs_wgs84) %>%
  sf::st_make_valid() %>%
  sf::st_intersection(us_cont_bds) %>%
  dplyr::rename(State = STATENAM) %>%
  dplyr::rename(Name = NHGISNAM) %>%
  dplyr::rename(Year = DECADE) %>%
  dplyr::select(Name, GISJOIN, State, Year)  

us_counties_1990 <- sf::st_read(paste0(yanelli_external, "spatial_data/nhgis0018_shape/nhgis0018_shapefile_tl2000_us_county_1990/US_county_1990.shp")) %>%
  sf::st_transform(crs_wgs84) %>%
  sf::st_make_valid() %>%
  sf::st_intersection(us_cont_bds) %>%
  dplyr::rename(State = STATENAM) %>%
  dplyr::rename(Name = NHGISNAM) %>%
  dplyr::rename(Year = DECADE) %>%
  dplyr::select(Name, GISJOIN, State, Year)  

us_counties_2000 <- sf::st_read(paste0(yanelli_external, "spatial_data/nhgis0018_shape/nhgis0018_shapefile_tl2000_us_county_2000/US_county_2000.shp")) %>%
  sf::st_transform(crs_wgs84) %>%
  sf::st_make_valid() %>% 
  sf::st_intersection(us_cont_bds) %>%
  dplyr::rename(State = STATENAM) %>%
  dplyr::rename(Name = NHGISNAM) %>%
  dplyr::rename(Year = DECADE) %>%
  dplyr::select(Name, GISJOIN, State, Year) 

us_counties_2010 <- sf::st_read(paste0(yanelli_external, "spatial_data/nhgis0018_shape/nhgis0018_shapefile_tl2010_us_county_2010/US_county_2010.shp")) %>%
  sf::st_transform(crs_wgs84) %>%
  sf::st_make_valid() %>%
  sf::st_intersection(us_cont_bds) %>%
  dplyr::rename(Name = NAME10) %>%
  dplyr::mutate(Year = 2010) %>% ## adding year column
  sf::st_join(us_states[,c("State")], largest = TRUE) %>% # add state column 
  dplyr::select(Name, GISJOIN, State, Year)  


### Bind all spatial files and estimate county area

us_counties <- rbind(us_counties_1970, us_counties_1980, us_counties_1990, us_counties_2000, us_counties_2010) %>%
  sf::st_transform(4326) %>%
  clean_names()
us_counties$area_km2 <- sf::st_area(us_counties)
us_counties$area_km2 <- units::set_units(us_counties$area_km2, km^2)
us_counties$area_km2 <- as.numeric(us_counties$area_km2)

us_counties <- us_counties %>%
  dplyr::select(gisjoin, year, state, name, area_km2, everything() ) %>%
  mutate(state = if_else(state == "District of Columbia", "District Of Columbia", state),
         name = if_else(name == "District Of Columbia", "District of Columbia", name))

class(us_counties)

#saveRDS(us_counties, paste0(yanelli_external, "spatial_data/updated_generated_files/geometries_counties.rds"))
```

## 3. Estimate population density

Estimate population density and merge the variable with all of the other covariates 

```{r}

us_counties_df <- us_counties
sf::st_geometry(us_counties_df) <- NULL # remove geometry 

pop_density <- left_join(data, us_counties_df, by = c("year", "gisjoin", "state")) %>%
  mutate(pop_density = tot_pop/area_km2) %>%
  dplyr::select(gisjoin, year, state, statefp, statenh, county, countyfp, countynh, name, area_km2, tot_pop, pop_density, everything())

class(pop_density)
summary(pop_density)


#saveRDS(pop_density, paste0(data.folder, "generated_data/data_indp_varib_4.rds"))
```

