---
title: "Matching emissions and census data"
author: "Yanelli Nunez"
date: "2/25/2021"
output:
 html_document:
    toc: TRUE
    toc_float: TRUE
    code_folding: hide
editor_options: 
  chunk_output_type: console
---

Objective:

- Combine emissions data 
- Estimate emissions change
- Merge emissions with covariates

Note: I commented out the line code for saving the generated dataset because if ran it will overwrite the already saved dataset



```{r include=FALSE}

rm(list=ls())

# 1a Declare root directory

project.folder <- paste0(print(here::here()),'/')

# add folder locations
source(paste0(project.folder,'0_00_create_folder_structure.R'))

# add file locations
#source(paste0(file.locations.folder,'file_locations.R'))

# load packages
source(paste0(packages.folder,'packages_to_load.R'))


```


## 1. Load data

The rds file named data_indp_varib_4 contains all of the independent variables 

```{r}

inde_vari <- readRDS(paste0(data.folder, "generated_data/data_indp_varib_4.rds")) %>%
  as.data.frame() %>%
  mutate(county = if_else(gisjoin == "G1200250", "Miami-Dade county", county), # rename Dade to Miami-Dade so it is consistent naming across the 4 years
         gisjoin = if_else(gisjoin == "G1200250", "G1200860", gisjoin),
         name = if_else(gisjoin == "G1200860", "Miami-Dade", name)) # rename Dade to Miami-Dade so it is consistent naming across the 4 years
        

summary(inde_vari)

class(inde_vari)

```


## 2. Pollutants data

- NOx from transport
- SO2 from energy  --> minimum value is zero
- NH3 from agriculture
- NOx from energy --> minimum value is zero
- SO2 from industry --> minimum value is zero
- OC from residential 
- NOx from commercial

- These datasets included a climate regions variable created by Jaime but I also created the same variable using his code in y_03_urbanicity_epa_regions.Rmd. So basically the variable is duplicated. 
- The data also includes year 2017, which we do not use in these analysis

```{r}


# NOx transp

nox_transp <- readRDS(paste0(data.folder, "emissions_data/emis_nox_tr_county_df.rds")) %>%
  clean_names() %>%
  mutate(year = as.numeric(year)) %>%
  mutate(state = if_else(state == "District of Columbia", "District Of Columbia", state)) # in the covariates dataframe the name is District 'Of' Columbia so I editing it here to match that format

summary(nox_transp)

# NH3 agric

nh3_agric <- readRDS(paste0(data.folder, "emissions_data/emis_nh3_agric_county_df.rds")) %>%
  clean_names() %>%
  mutate(year = as.numeric(year)) %>%
  mutate(state = if_else(state == "District of Columbia", "District Of Columbia", state))

summary(nh3_agric)
which(is.na(nh3_agric))

# SO2 energ

so2_ener <- readRDS(paste0(data.folder, "emissions_data/emis_so2_ener_county_df.rds")) %>%
  clean_names() %>%
  mutate(year = as.numeric(year)) %>%
  mutate(state = if_else(state == "District of Columbia", "District Of Columbia", state))

# Nox energy

nox_energ <- readRDS(paste0(data.folder, "emissions_data/emis_nox_ener_county_df.rds")) %>%
  clean_names() %>%
  mutate(year = as.numeric(year)) %>%
  mutate(state = if_else(state == "District of Columbia", "District Of Columbia", state))

summary(nox_energ)

# SO2 industry

so2_indus <- readRDS(paste0(data.folder, "emissions_data/emis_so2_industry_county_df.rds")) %>%
  clean_names() %>%
  mutate(year = as.numeric(year)) %>%
  mutate(state = if_else(state == "District of Columbia", "District Of Columbia", state))

# oc residential

oc_resid <- readRDS(paste0(data.folder, "emissions_data/emis_oc_residential_county_df.rds")) %>%
  clean_names() %>%
  mutate(year = as.numeric(year)) %>%
  mutate(state = if_else(state == "District of Columbia", "District Of Columbia", state))

# nox commercial

nox_commer <- readRDS(paste0(data.folder, "emissions_data/emis_nox_commercial_county_df.rds")) %>%
  clean_names() %>%
  mutate(year = as.numeric(year)) %>%
  mutate(state = if_else(state == "District of Columbia", "District Of Columbia", state))


# joing all datasets
poll <- full_join(nh3_agric, so2_ener, by = c("name", "gisjoin", "state", "year", "climate_region")) %>%
  full_join(., nox_transp) %>%
  full_join(., nox_energ) %>%
  full_join(., so2_indus) %>%
  full_join(., oc_resid) %>%
  full_join(., nox_commer) %>%
  dplyr::select(gisjoin, year, state, name, climate_region, everything())  %>%
  mutate(name = if_else(gisjoin == "G1200250", "Miami-Dade", name), # rename Dade to Miami-Dade so it is consistent naming across the study period
         gisjoin = if_else(gisjoin == "G1200250", "G1200860", gisjoin))

class(poll)
summary(poll)

# save files with all emission year 1970 to 2017
# saveRDS(poll, paste0(data.folder, "generated_data/emissions_all_data.rds"))

```

### 2a. Estimate emission differences and relat change (years 1970 to 2010)

Notes: 

- Relative change was estimated as a percentage of the initial concentration
- Relative change was estimated as a percentage of the initial concentration, thus, for variables that have a starting concentration of zero we get undefined values(number/zero --> undefined).

- ***Beware, the variable 'name' seems to have discrepancies in writing and the relative matching to gisjoin. This is a not to myself to not use this variable.

```{r}
  
  
poll_reltv_chg <- poll %>%
  filter(year != 2017) %>%
  mutate(gisjoin = as.factor(gisjoin)) %>%
  group_by(state, gisjoin) %>%
 arrange(., year, .by_group = TRUE) %>%
    mutate(nox_trans_diff = (nox_trans - dplyr::lag(nox_trans)),
           nox_trans_diff = dplyr::lead(nox_trans_diff), # 'lead' to assign difference value to earlier year
           nh3_agric_diff = (nh3_agric - dplyr::lag(nh3_agric)),
           nh3_agric_diff = dplyr::lead(nh3_agric_diff),
           so2_energy_diff = (so2_ener - dplyr::lag(so2_ener)),
           so2_energy_diff = dplyr::lead(so2_energy_diff),
           nox_energy_diff = (nox_ener - dplyr::lag(nox_ener)),
           nox_energy_diff = dplyr::lead(nox_energy_diff),
           so2_indus_diff = (so2_industry - dplyr::lag(so2_industry)),
           so2_indus_diff = dplyr::lead(so2_indus_diff),
           oc_resd_diff = (oc_residential - dplyr::lag(oc_residential)),
           oc_resd_diff = dplyr::lead(oc_resd_diff),
           nox_commer_diff = (nox_commercial - dplyr::lag(nox_commercial)),
           nox_commer_diff = dplyr::lead(nox_commer_diff)) %>% 
  ungroup() %>%
  mutate(nox_trans_relat_chg = nox_trans_diff/nox_trans*100, # relative change
         nh3_agric_relat_chg = nh3_agric_diff/nh3_agric*100,
         oc_resid_relat_chg = oc_resd_diff/oc_residential*100,
         nox_commer_relat_chg = nox_commer_diff/nox_commercial*100, 
         so2_energy_relat_chg = so2_energy_diff/so2_ener*100,
         nox_energy_relat_chg = nox_energy_diff/nox_ener*100, 
         so2_indus_relat_chg = so2_indus_diff/so2_industry*100) %>%
   as.data.frame() %>%
   mutate(name = if_else(state == "District Of Columbia", "District of Columbia", name),
          year = as.factor(year))

summary(poll_reltv_chg)
class(poll_reltv_chg)

# save file
 #saveRDS(poll_reltv_chg, paste0(data.folder, "generated_data/emissions_diff_reltv_chg.rds"))

```


### 3c. Bind independent variables w/emissions data

```{r}
full_data <- inde_vari %>%
  full_join(., poll_reltv_chg, by = c("year", "gisjoin", "state", "name", "climate_region")) %>%
 dplyr::select(gisjoin, year, state, statefp, statenh, county, countyfp, countynh, name, urbanicity, climate_region, epa_reg, area_km2, tot_pop, pop_density, everything())


summary(full_data)
class(full_data)

# saveRDS(full_data, paste0(data.folder, "generated_data/master_data.rds"))
```


## 4. Visualizations

A couple of counties are missing emissions data. They are small areas and peninsulas 

```{r}
missing_emissions <- full_data %>%
dplyr::select(gisjoin, year, state, county, tot_pop, area_km2, pop_density, 48:68) %>%
  filter(year != 2010)

summary(missing_emissions)
```

Notes:
- Coffey County, Kansas has a very huge relative increase in SO2 from Energy
```{r}

# Organize data for figures 

data_fig <- full_data %>%
  dplyr::select(year, state, county, 48:68) %>%
  gather(variable, estimate, 4:24) %>%
  unite(ID, c("state", "county"), sep = "_", remove = FALSE) %>%
  mutate(ID = as.factor(ID),
         variable = as.factor(variable))

  
# re order factor 
data_fig$variable <- factor(data_fig$variable, levels = c(
                "nox_trans",
                "nox_trans_diff",
                "nox_trans_relat_chg",
                "nox_trans_chg_med",
                "nh3_agric",
                "nh3_agric_diff",
                "nh3_agric_relat_chg",
                "nh3_agric_chg_med",
                "so2_ener",
                "so2_energy_diff",
                "so2_energy_relat_chg",
                "so2_ener_chg_med",
                "oc_residential",
                "oc_resd_diff",
                "oc_resid_relat_chg",
                "nox_commercial",
                "nox_commer_diff",
                "nox_commer_relat_chg",
                "so2_industry",
                "so2_indus_diff",
                "nox_ener",
                "nox_energy_diff"))

# variable distribution 

# density plot
data_fig %>%
  ggplot(aes(estimate)) +
  facet_wrap(~variable, scales = "free") +
  geom_density(aes(y = ..ndensity..))

# box plot
 data_fig %>%
  ggplot(aes(y = estimate)) +
  facet_wrap(~variable, scales = "free") +
 geom_boxplot(aes(x=variable)) +
  # geom_point(aes(x = variable, color = ID), alpha = 0.5, size = 0.5) +
theme(legend.position = "none") 
 

#ggplotly(box_plots, tooltip = c("ID", "estimate"))

```

