---
title: "Preping energy sector data"
author: "Yanelli Nunez"
date: "11/24/2021"
output:
 html_document:
    toc: TRUE
    toc_float: TRUE
    code_folding: hide
editor_options: 
  chunk_output_type: console
---

Objective:  

- Remove the extreme values from the SO2 and NOx energy sector 
- Convert NaN to zeros and remove the few infinity values in SO2 industry

Note: I commented out the line code for saving the generated dataset because if ran it will overwrite the already saved dataset


```{r include=FALSE}

rm(list=ls())

# 1a Declare root directory

project.folder <- paste0(print(here::here()),'/')

# add folder locations
source(paste0(project.folder,'0_00_create_folder_structure.R'))

# add file locations
source(paste0(file.locations.folder,'external_file_locations.R'))

# load packages
source(paste0(packages.folder,'packages_to_load.R'))

```

## 1. Load data

```{r}

# Data
data <- readRDS(paste0(data.folder, "generated_data/master_data.rds"))
class(data)
summary(data)


data_filt_no10 <- data %>%
  filter(year != 2010)

```

## 2. Visualizing Energy sector

There are a couple of very extreme values because some counties have a very small number that changes to a very large number. These few extreme values completely drive the health models and lead to non-reliable extreme effect estimates

```{r}

# SO2 energy 
hist(data$so2_energy_relat_chg, breaks = 10000, xlim=c(-100,18))

summary(data$so2_energy_relat_chg)
mean(data$so2_energy_relat_chg, na.rm = TRUE)
sd(data$so2_energy_relat_chg, na.rm = TRUE)
# values above the 95th percentile are huge
quantile(data$so2_energy_relat_chg, seq(0,1,0.01), na.rm=TRUE) 


#NOx energy

hist(data$nox_energy_relat_chg, breaks = 10000, xlim=c(-100,18))

summary(data$nox_energy_relat_chg)
mean(data$nox_energy_relat_chg, na.rm = TRUE)
sd(data$so2_energy_relat_chg, na.rm = TRUE)
# values above the 95th percentile are huge
quantile(data$nox_energy_relat_chg, seq(0,1,0.01), na.rm=TRUE) 

```

## 3. Energy sector

- We selected the 95th percentile of the data to avoid extremely large values that are generated when a county's reference year has an extremely low emission flux.
- These are the data that I used for the models for the energy sector

Notes:

Excluding the top 5 percentile values removes 3754 observations but 3133 are NA's (mostly for year 2010 because for this year we do not estimate the relative change). 621 observations correspond to the 5% values of the 5 percentile 


### 2a. SO2 energy 


```{r}

# filter observations that are below the 95th percentile
data_so2_energy <- data_filt_no10 %>%
  dplyr::filter(so2_energy_relat_chg <= quantile(data$so2_energy_relat_chg, 0.95, na.rm=TRUE))

extreme_so2_energy <- data_filt_no10 %>%
  dplyr::filter(so2_energy_relat_chg <= quantile(data$so2_energy_relat_chg, 0.95, na.rm=TRUE))

summary(data_so2_energy)
# much more reasonable
summary(data_so2_energy$so2_energy_relat_chg)
sd(data_so2_energy$so2_energy_relat_chg)
plot(density(data_so2_energy$so2_energy_relat_chg, na.rm = TRUE))

# save data
#saveRDS(data_so2_energy, paste0(data.folder, "generated_data/master_so2_energy.rds"))

```

### 2b. NOx energy

```{r}

# filter observations that are below the 95th percentile
# Note that the code below removes the 2010 observations which have NA for the relative diff variables
data_nox_energy <- data_filt_no10 %>%
  dplyr::filter(nox_energy_relat_chg <= quantile(data$nox_energy_relat_chg, 0.95, na.rm=TRUE ))

summary(data_nox_energy)
# much more reasonable
summary(data_nox_energy$nox_energy_relat_chg)
sd(data_nox_energy$nox_energy_relat_chg)

# save data
#saveRDS(data_nox_energy, paste0(data.folder, "generated_data/master_nox_energy.rds"))

```

## 3. SO2 from Industry

- Counties that have zero in two consecutive years have a NaN value for the relative change (because 0-0/0 --> NaN). So I will replace this NaN with zero because basically the change in emissions, for the counties that have consecutive zeros, is zero!

*Dealing with infinity values*
- There is no best way to deal with infinity values so I removed all *infinity values*
- The good thing is that there are only have 5 counties that have infinity values because they went from 0 in 2000 to nonzero in 2010 (nonzero - 0 / 0 --> infinity)


Below are notes I was making for myself. Feel free to ignore 
- 640 observations have zeros for SO2 industry 
- Only 11 counties have zero at 1970
- There is 152 counties that have zero in 1980-2000. 
- The 11 counties with zero in 1970 are within the 152 counties with zero in the later years. This means that 141 counties went from non-zero to zero from 1970 to 1980 but after that they stay as zeros
- There are 147 counties with zero in 2010 which means that 5 counties went from zero in 2000 to non-zero in 2010. 


```{r}

##### THIS IS WHAT I SHOULD HAVE DONE TO REMOVE ONLY THE OBSERVATIONS WITH INFINITY VALUES RATHER THAT COMPLETELY
##### REMOVING THE COUNTIES FROM THE DATASET. I SHOULD ONLY REMOVE THE OBSERVATIONS WITH INFINITY VALUES.
##### ONCE I HEAR BACK FROM REVIEWERS I NEED TO RE-RUN THE INDUSTRY ANALYSIS USING THE DATASET THAT ONLY EXCLUDES OBSERVATIONS RATHER 
##### THAT REMOVING THE COUNTIES FROM THE WHOLE ANALYSIS

so2_review <- data_filt_no10 %>%
 mutate_at("so2_indus_relat_chg",  ~replace(., is.nan(.), 0)) %>% ##  replace NaN values with zero
  dplyr::filter(!is.infinite(so2_indus_relat_chg)) ## remove observations (NOT COUNTIES) with infinity values

# save file 
# okay so now the dataset that I have for industry is correct
#saveRDS(so2_review, paste0(data.folder, "generated_data/master_so2_industry.rds"))

################# From here and below it was a shit show that i don't know why I did all of this nonsense #######

### I definitely went the long way here... 
# view distribution
histogram(data$so2_industry)

sum(is.nan(data_filt_no10$so2_indus_relat_chg)) # 462 NaN

# Select only counties with zero or
so2_industry <- data %>%
  filter(so2_industry == 0)

summary(so2_industry)

# number of counties with zero per year
so2_industry %>%
  group_by(year) %>%
  summarise(n())

# create a unite ID per county/state
view_zeros <- so2_industry %>%
   unite(ID_county, c("state", "county")) %>%
  ungroup()

yr1970 <- view_zeros %>%
  filter(year == 1970) %>%
distinct(ID_county)

yr1980 <- view_zeros %>%
  filter(year == 1980) %>%
distinct(ID_county) 

yr1990 <- view_zeros %>%
  filter(year == 1990) %>%
distinct(ID_county) 

yr2000 <- view_zeros %>%
  filter(year == 2000) %>%
distinct(ID_county) 

yr2010 <- view_zeros %>%
  filter(year == 2010) %>%
distinct(ID_county) 

# counties with zero in 1980-2000 overlap 
zero_counties <- dplyr::full_join(yr1970, yr1980) %>%
  full_join(., yr1990) %>%
full_join(., yr2000) 
 # full_join(., yr2010) %>%
#separate(ID_county, c("state", "county"), sep = "_")

# counties with infinity values 
infinity_counties <- anti_join(zero_counties, yr2010) %>%
separate(ID_county, c("state", "county"), sep = "_")

# remove counties with infinity values and replace NaN with "0"
so2_indus_new <- anti_join(data, infinity_counties) %>%
 mutate_at("so2_indus_relat_chg",  ~replace(., is.nan(.), 0)) %>%
  dplyr::filter(year != "2010")

summary(so2_indus_new)

sum(is.nan(so2_indus_new$so2_indus_relat_chg))
sum(is.na(so2_indus_new$so2_indus_relat_chg))


summary(so2_indus_new)
summary(data)


```

