---
title: "Matching emissions and census data"
author: "Yanelli Nunez"
date: "2/25/2021"
output:
 html_document:
    toc: TRUE
    toc_float: TRUE
    code_folding: hide
editor_options: 
  chunk_output_type: console
---

Objective:

Estimate VOC emission change and relative change and add these data to the ongoing datasets

```{r include=FALSE}

rm(list=ls())

# 1a Declare root directory

project.folder <- paste0(print(here::here()),'/')

# add folder locations
source(paste0(project.folder,'0_00_create_folder_structure.R'))

# add file locations
#source(paste0(file.locations.folder,'file_locations.R'))

# load packages
source(paste0(packages.folder,'packages_to_load.R'))


```


## 1. Load data


```{r}

# emissions data only
emi_data <- readRDS(paste0(data.folder, "generated_data/emissions_all_data.rds")) 

# independent variables joined with emissions data ("master" dataset)
master <- readRDS(paste0(data.folder, "generated_data/master_data.rds")) 

```


## 2. Pollutants data

Here we are only preparing the VOC data. Note that there are multiple VOCs, so we summed up all VOCs into a total VOC value and worked with that total value

- VOC from energy --> min 0.00083

Below info is just for references or reminder of where the other pollutants stand

- NOx from transport
- SO2 from energy  --> minimum value is zero
- NH3 from agriculture
- NOx from energy --> minimum value is zero
- SO2 from industry --> minimum value is zero
- OC from residential 
- NOx from commercial



```{r}

# VOC energy

voc_energ <- readRDS(paste0(data.folder, "emissions_data/us_counties_vocs.rds")) %>%
  clean_names() %>%
  dplyr::select(1:5, voc_total) %>%
  mutate(year = as.numeric(year)) %>%
  mutate(state = if_else(state == "District of Columbia", "District Of Columbia", state)) %>% # in the covariates dataframe the name is District 'Of' Columbia so I am  editing it here to match that format
  mutate(name = if_else(gisjoin == "G1200250", "Miami-Dade", name), # rename Dade to Miami-Dade so it is consistent naming across the study period
         gisjoin = if_else(gisjoin == "G1200250", "G1200860", gisjoin)) %>%
  rename(voc_energ = voc_total)


summary(voc_energ)
hist(voc_energ$voc_energ)


# add VOC data to the emissions dataset

pollu_data <- left_join(emi_data, voc_energ) 

class(pollu_data)
summary(pollu_data)

# save files with all emission year 1970 to 2017
# saveRDS(pollu_data, paste0(data.folder, "generated_data/emissions_all_data.rds"))

```

### 2a. Estimate emission differences and relat change (years 1970 to 2010)

Notes: 

- Relative change was estimated as a percentage of the initial concentration
- Relative change was estimated as a percentage of the initial concentration, thus, for variables that have a starting concentration of zero we get undefined values(number/zero --> undefined).

- ***Beware, the variable 'name' seems to have discrepancies in writing and the relative matching to gisjoin. This is a note to myself to not use this variable.

```{r}
  
  
voc_reltv_chg <- voc_energ %>%
  filter(year != 2017) %>%
  mutate(gisjoin = as.factor(gisjoin)) %>%
  group_by(state, gisjoin) %>%
 arrange(., year, .by_group = TRUE) %>%
    mutate(voc_energ_diff = (voc_energ - dplyr::lag(voc_energ)),
           voc_energ_diff = dplyr::lead(voc_energ_diff)) %>%  # 'lead' to assign difference value to earlier year
  ungroup() %>%
  mutate(voc_energ_relat_chg = voc_energ_diff/voc_energ*100) %>% # relative change
   as.data.frame() %>%
   mutate(name = if_else(state == "District Of Columbia", "District of Columbia", name),
          year = as.factor(year))

summary(voc_reltv_chg)
class(poll_reltv_chg)
hist(voc_reltv_chg$voc_energ_relat_chg)

# read ongoing data
poll_reltv_chg <- readRDS(paste0(data.folder, "generated_data/emissions_diff_reltv_chg.rds"))

# add voc data to the ongoing dataset of pollutatns
reltv_chg_join <- left_join(poll_reltv_chg, voc_reltv_chg)

# save file
#saveRDS(reltv_chg_join, paste0(data.folder, "generated_data/emissions_diff_reltv_chg.rds"))

```


### 3c. Add VOC data to previous master dataset 

```{r}

master_update_voc <- left_join(master, voc_reltv_chg)

summary(master_update_voc)
class(master_update_voc)

colnames(master_update_voc)

#saveRDS(master_update_voc, paste0(data.folder, "generated_data/master_data.rds"))

```


## 4. Visualizations

A couple of counties are missing emissions data. They are small areas and peninsulas 

```{r}

missing_emissions <- master_update_voc %>%
dplyr::select(gisjoin, year, state, county, tot_pop, area_km2, pop_density, 48:71) %>%
  filter(year != 2010)

colnames(missing_emissions)

summary(missing_emissions)

```

Notes:


```{r}

# Organize data for figures 

data_fig <- master_update_voc %>%
  dplyr::select(year, state, county, 48:71) %>%
  gather(variable, estimate, 4:27) %>%
  unite(ID, c("state", "county"), sep = "_", remove = FALSE) %>%
  mutate(ID = as.factor(ID),
         variable = as.factor(variable))

colnames(data_fig)

  
# re order factor 
data_fig$variable <- factor(data_fig$variable, levels = c(
                "nox_trans",
                "nox_trans_diff",
                "nox_trans_relat_chg",
                
                "nh3_agric",
                "nh3_agric_diff",
                "nh3_agric_relat_chg",
            
                "so2_ener",
                "so2_energy_diff",
                "so2_energy_relat_chg",
            
                "oc_residential",
                "oc_resd_diff",
                "oc_resid_relat_chg",
                
                "nox_commercial",
                "nox_commer_diff",
                "nox_commer_relat_chg",
                
                "so2_industry",
                "so2_indus_diff",
                "so2_indus_relat_chg",
                
                "nox_ener",
                "nox_energy_diff",
                "nox_energy_relat_chg",
                
                "voc_energ",
                "voc_energ_diff",
                "voc_energ_relat_chg"))

# variable distribution 

# density plot
data_fig %>%
  ggplot(aes(estimate)) +
  facet_wrap(~variable, scales = "free") +
  geom_density(aes(y = ..ndensity..))

# box plot
 data_fig %>%
  ggplot(aes(y = estimate)) +
  facet_wrap(~variable, scales = "free") +
 geom_boxplot(aes(x=variable)) +
  # geom_point(aes(x = variable, color = ID), alpha = 0.5, size = 0.5) +
theme(legend.position = "none") 
 
 
 

#ggplotly(box_plots, tooltip = c("ID", "estimate"))

```

